---
title: "Language coding"
author: "Rémi Anselme"
date: "2023-10-26"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 4
    toc_float: yes
    df_print: paged
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: '3'
  html_notebook:
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
`%>%` <- magrittr::`%>%`
```


# Data preparation - From Aleksandra Ćwiek

```{r pckgs}
#library(tidyverse)

# Set the script's path as working directory
parentfolder = rstudioapi::getActiveDocumentContext()$path 
#setwd(dirname(parentfolder))
parentfolder <- getwd()

data <- paste0(parentfolder, '/data/')
```

Load the data

```{r load data}
web <- read.csv(paste0(data, "r_l_web.csv"), sep = ",")
```

## Web dataset cleaning

Before cleaning we have `r length(unique(web$l1))` languages in the column `l1`.  
Before cleaning we have `r length(unique(web$l2))` languages in the column `l2`.  
Before cleaning we have `r length(unique(web$firstlanguage))` languages in the column `firstlanguage`.  
Before cleaning we have `r length(unique(web$foreignlanguages))` combinations for languages in the column `foreignlanguages`.  
For the data coding, I won't work with `firstlanguage` column and `foreignlanguages`.

```{r}
#I looked languages that have not a language name (some are Aleksandra trials)
to_del <- c("b","q","3","hi") 

web <- web %>% 
  dplyr::filter(!(l1 %in% to_del) & !is.na(l1))

languages <- web %>% 
                dplyr::mutate(L1 = stringr::str_extract_all(l1, "\\b[:alpha:]+\\b"),
                             L2 = stringr::str_extract_all(l2, "\\b[:alpha:]+\\b")) 

#Some of the languages have been replaced based on Aleksandra's preparation script
#The objective here is to have a maximum of languages to code

languages_data <- unique(c(unlist(languages$L1),unlist(languages$L2))) %>% 
                as.data.frame() %>% 
                dplyr::rename(Languages = 1) %>% 
                dplyr::mutate(Languages = as.character(Languages)) %>% 
                dplyr::distinct() %>% 
                dplyr::filter(!is.na(Languages) & Languages != "a" & Languages != "little" & Languages != "bit") %>% 
                dplyr::mutate(Languages = stringr::str_replace(Languages, 'isizulu', 'zulu'),
                              Languages = stringr::str_replace(Languages, 'turkce', 'turkish'),
                              Languages = stringr::str_replace(Languages, 'albanianja', 'albanian'),
                              Languages = stringr::str_replace(Languages, 'español', 'spanish'),
                              Languages = stringr::str_replace(Languages, 'italia', 'italian'),                  
                              Languages = stringr::str_replace(Languages, 'italiann', 'italian'),
                              Languages = stringr::str_replace(Languages, 'ancient', 'ancient greek'),
                              Languages = stringr::str_replace(Languages, 'potuguese', 'portuguese'),
                              Languages = stringr::str_replace(Languages, 'spanich', 'spanish'),
                              Languages = stringr::str_replace(Languages, 'calatan', 'catalan'),
                              Languages = stringr::str_replace(Languages, 'rumanian', 'romanian'),
                              Languages = stringr::str_replace(Languages, 'estniska', 'estonian'),
                              Languages = stringr::str_replace(Languages, 'dari', 'farsi'),
                              Languages = stringr::str_replace(Languages, 'sami', 'saami')) %>% 
                dplyr::distinct() %>% 
                dplyr::arrange(Languages) %>% 
  dplyr::mutate(glottocode = as.character(NA),
                r_l_distinction = as.integer(NA),
                trill_real = as.integer(NA),
                trill_occ = as.integer(NA))

```

In total, we have `r length(unique(languages_data$Languages))` languages to check.


```{r}
unique(languages_data$Languages)
```
# Coding

## Langue - Exemple

```{r}
# languages_data <- languages_data %>% 
#     dplyr::mutate(glottocode = ifelse(Languages=="LANGUAGE","GLOTTOCODE",glottocode))
```

Phoible :   

Wikipedia :   

Glottologue :  

Autres sources :  

```{r}
# languages_data <- languages_data %>% 
#     dplyr::mutate(r_l_distinction = ifelse(Languages=="LANGUE",10,r_l_distinction))
# 
# languages_data <- languages_data %>% 
#     dplyr::mutate(trill_real = ifelse(Languages=="LANGUE",10,trill_real))
# 
# languages_data <- languages_data %>% 
#     dplyr::mutate(trill_occ = ifelse(Languages=="LANGUE",10,trill_occ))
```

## NGT

```{r}
languages_data <- languages_data %>% 
     dplyr::mutate(glottocode = ifelse(Languages=="NGT","dutc1253",glottocode))
```

Phoible : NA  

Wikipedia :  
https://en.wikipedia.org/wiki/Dutch_Sign_Language

Glottologue :    
https://glottolog.org/resource/languoid/id/dutc1253

Autres sources : NA  

```{r}
languages_data <- languages_data %>% 
     dplyr::mutate(r_l_distinction = ifelse(Languages=="NGT",as.integer(NA),r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="NGT",as.integer(NA),trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="NGT",as.integer(NA),trill_occ))
```

## afrikaans

```{r}
languages_data <- languages_data %>% 
    dplyr::mutate(glottocode = ifelse(Languages=="afrikaans","afri1274",glottocode))
```

Phoible :   
https://phoible.org/inventories/view/1395  
      - r allophone r  ; contrast r/l

Donaldson, Bruce C. 1993. A grammar of Afrikaans. Mouton de Gruyter.  
  - Page 15 : 1.4.12. The consonant r [r]  
              Examples: rot [rot] 'rat', raar [ra:r] 'queer*, verder [f^rdar] 'further'.  
              R is trilled in Afrikaans. In various areas of the Cape uvular r, both [R] and [ʁ], occurs and in the north is regarded as a shibboleth of a "Capie" (see               footnote 9). It is called gebryde r, e.g. Hy bry, want hy kom van Malmesbury se wêreld αf 'He has a uvular r because he comes from the Malmesbury area'.  
  - Footnote 9 Page 6 : If one wants to parody a Transvaler, one uses this sound, whereas the countryside of the Western Cape is parodied by using the uvular r and raising ee and oo to [i] and [u], e.g. lees [lis] 'to read', brood [bRut]'bread'.  
  
Wikipedia :   
https://en.wikipedia.org/wiki/Afrikaans  
  - Rhotic r  
  - /r/ is usually an alveolar trill [r] or tap [ɾ].[109] In some parts of the former Cape Province, it is realized uvularly, either as a trill [ʀ] or a fricative [ʁ].[110]
    [109] - Lass, Roger (1987), "Intradiphthongal Dependencies", in Anderson, John; Durand, Jacques (eds.), Explorations in Dependency Phonology, Dordrecht: Foris Publications Holland, pp. 109–131, ISBN 90-6765-297-0  
    [110] - Donaldson (1993), p. 15.  
    
Glottologue :  
https://glottolog.org/resource/languoid/id/afri1274  

Autres sources :  

Wissing, D. (2020). Afrikaans. Journal of the International Phonetic Association, 50(1), 127-140. doi:10.1017/S0025100318000269  
  - Page 130 : Trill r  ʀ
  - Page 131 : The alveolar trill [r] varies with the uvular trill [ʀ], the latter found notably in the Boland region of the Western Cape, but also quite frequently in other areas of SA (Pienaar 2017 provides an overview and mentions relevant literature; see also Ribbens-Klein 2016).  
  - Page 132 : In colloquial speech, the voiced alveolar trill /r/ tends to be omitted in syllable-final position. This happens mostly in unstressed syllables, and regularly in the prefix ver-, as in /fərˈkeːrd/ > [fəkiərt] verkeerd ‘wrong’. In word-final position, as in /lɛkər/ > [lækə] lekker ‘nice’, /r/-deletion is quite common across the board, especially in function words like daar, maar, hier, vir: /daːr/ > [daː] daar ‘there’, /maːr/ > [maː]7 maar ‘but’, /foːr/ > [fuə] voor ‘in front of’, and /fər/ > [fə] vir8 ‘for’. This phenomena is widely present in Malayan Afrikaans. Kotz ́e (1983) mentions similar examples, but adds to these the loss of /r/, even in complex codas, e.g. [fəket] verkeerd ‘wrong’, [bɔːs] bors ‘breast’, and [kənəs] kinders ‘children’ –transcriptions are Kotz ́e’s.  

```{r}
languages_data <- languages_data %>% 
    dplyr::mutate(r_l_distinction = ifelse(Languages=="afrikaans",1,r_l_distinction))

languages_data <- languages_data %>% 
    dplyr::mutate(trill_real = ifelse(Languages=="afrikaans",1,trill_real))

languages_data <- languages_data %>% 
    dplyr::mutate(trill_occ = ifelse(Languages=="afrikaans",1,trill_occ))
```

## albanian

```{r}
languages_data <- languages_data %>% 
     dplyr::mutate(glottocode = ifelse(Languages=="albanian","tosk1239",glottocode))
```

Phoible :   
https://phoible.org/languages/tosk1239  
  - SPA r allophones ɾ r  ; contrast r/l
  - UPSID r̪|r pas d'allophones  ; contrast 	r̪|r/l
  - EA r pas d'allophones  ; contrast  r/l

Newmark, Leonard. 1957. Structural Grammar of Albanian. (International Journal of American Linguistics, 23.) Bloomington: Indiana University.  

Klippenstein, Rachel. 2010. Word-initial consonant clusters in Albanian. In Lesho, Marivic and Smith, Bridget J. and Campbell-Kibler, Kathryn and Culicover, Peter W. (eds.), Working Papers in Linguistics, 10–32.  

Camaj, Martin and Fox, Leonard. 1984. Albanian grammar: with exercises, chrestomathy and glossaries. Otto Harrassowitz Verlag.  

Wikipedia :    
https://en.wikipedia.org/wiki/Tosk_Albanian  

https://en.wikipedia.org/wiki/Albanian_language#Phonology  
  - Flap 			ɾ 					  
    Trill 			r  
  - The contrast between flapped r and trilled rr is the almost the same as in Spanish or Armenian. However, in most of the dialects, as also in standard Albanian, the single "r" changes from an alveolar flap /ɾ/ to an alveolar approximant [ɹ].  
  
Glottologue :   
https://glottolog.org/resource/languoid/id/tosk1239  

Autres sources :  

Coretta, S., Riverin-Coutlée, J., Kapia, E., & Nichols, S. (2022). Northern Tosk Albanian. Journal of the International Phonetic Association, 1-23. doi:10.1017/S0025100322000044  
  - Page 3 : Trill r  
             Flap ɽ  
  - Page 8 :  Rhotics  
              Northern Tosk contrasts two rhotics, written as ⟨r⟩ and ⟨rr⟩, both of which can occur word initially, medially and finally. While our speakers always realised ⟨rr⟩ as an alveolar trill [r], ⟨r⟩ was characterised by a great deal of allophonic variation, as also previously noted by Belluscio (Reference Belluscio, Gesuato and Grazia Busà2014). Among the range of realisations for ⟨r⟩, we found retroflex flaps [ɽ] and retroflex approximants [ɻ]. These seem to be in complementary distribution, with [ɻ] tending to appear in coda position, although for some speakers and in some words, word-initial /r/ was also realised as a retroflex approximant [ɻ]. Impressionistic studies and some case-study analyses suggest that /r/ may increasingly be realised as [ɻ] or variants thereof in different varieties of Albanian (Hysenaj Reference Hysenaj2009, Jubani-Bengu Reference Jubani-Bengu2012, Belluscio Reference Belluscio2016). Figure 8 shows illustrative examples of the three main rhotic allophones. The limited set of contexts in the words collected does not allow us to make definitive generalisations about the distribution of these allophones and future work is warranted on this matter. We have chosen to use /r/ for ⟨rr⟩ and /ɽ/ for ⟨r⟩, though as noted above, /ɽ/ is frequently realised as [ɻ].         


```{r}
languages_data <- languages_data %>% 
     dplyr::mutate(r_l_distinction = ifelse(Languages=="albanian",1,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="albanian",1,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="albanian",1,trill_occ))
```

## ancient greek

```{r}
languages_data <- languages_data %>% 
     dplyr::mutate(glottocode = ifelse(Languages=="ancient greek","anci1242",glottocode))
```

Phoible : NA  

Wikipedia :   
https://en.wikipedia.org/wiki/Ancient_Greek  
  - Trill 		ρ r
  - /r/ was probably voiceless when word-initial and geminated (written ῥ and ῥῥ).
  => il y a un contrast entre un r et un l
  
Glottologue :  
https://glottolog.org/resource/languoid/id/anci1242  

van Emde Boas, Evert and Rijksbaron, Albert and Huitink, Luuk and de Bakker, Mathieu. 2019. The Cambridge Grammar of Classical Greek. Cambridge: Cambridge University Press.
  - Page 11 : p [r] \rho\epsilon\omega [r^héo:] flow rhyme(rolling r)
  - Page 12 : The category of resonants consists of the nasal consonants \micro and v (and in certain cases \gamma) and the so-called‘liquids’ \lambda and \rho.
  - Page 13 : Liquids: two Greek letters belong to the class of liquids - \lambda and \rho: - a lateral consonant, \lambda [l]: air escapes along the sides of the tongue, but not through the middle of the mouth; - a rhotic consonant (or ‘tremulant’), \rho [r]: this was pronounced as a rolling, ‘alveolar trill’ sound, with the tongue vibrating against the gums. At the beginning of a word (and in somecases in the middle of a word), this sound was pronouncedwith aspiration (hence word-initial rho is always written \rho(avec diacritic au dessus])).

Autres sources : NA 

```{r}
languages_data <- languages_data %>% 
     dplyr::mutate(r_l_distinction = ifelse(Languages=="ancient greek",1,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="ancient greek",1,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="ancient greek",1,trill_occ))
```

## arabic

```{r}
languages_data <- languages_data %>% 
     dplyr::mutate(glottocode = ifelse(Languages=="arabic","stan1318",glottocode))
```

Phoible :   
https://phoible.org/languages/stan1318
  - r allophone r ; contrast r/l
  
Wikipedia :   
https://en.wikipedia.org/wiki/Modern_Standard_Arabic
  - Trill 			r ⟨ر⟩

https://en.wikipedia.org/wiki/Arabic_phonology
  - Trill 			r^[l]
  - [l] Emphatic [rˤ] exists in Northwestern African pronunciations and in Egypt when accompanied by /a/ or /u/ and plain when accompanied by /i/ or /j/; in closed syllables, then it is plain when the first preceding voweled consonant has /i/ or if /j/ is present, but emphatic if the first preceding voweled letter is accompanied by /a/ or /u/. The trill /r/ is sometimes reduced to a single vibration when single, but it remains potentially a trill, not a flap [ɾ]: the pronunciation of this single trill is between a trill [r] and a flap [ɾ]. ⟨r⟩ is in free variation between a trill [r] and a flap [ɾ] in Egypt and the Levant.
  
Glottologue :  
https://glottolog.org/resource/languoid/id/stan1318

Ryding, Karin C. 2005. A reference grammar of modern standard Arabic. New York: Cambridge Univ. Press. xxv+708pp. (Includes bibliographical references (p. 691-699) and index.)
  - Page 13 : Flaps r
  - Page 14 : raa (r) (symbole arabe) voiced alveolar flap or trill: as /r/ in Italian or Spanish; a good example in English is to pronounce the word "very" as "veddy"
  
Autres sources :

Thelwall, R., & Sa'Adeddin, M. (1990). Arabic. Journal of the International Phonetic Association, 20(2), 37-39. doi:10.1017/S0025100300004266
  - Page 37 : Trill r
  
Bahrani, N., & Modarresi Ghavami, G. (2021). Khuzestani Arabic. Journal of the International Phonetic Association, 51(2), 299-313. doi:10.1017/S0025100319000203 [Pas mis au propre]
  - Page 301 : Tap 
               R /Ri ˘l/ ﺭﻳﻞ foot 
               R ≥ /R ≥a ˘d/ [R ≥A˘d] ﺭﺍﺩ (he) wanted
  - Page 303 : Sonorant consonants include /m n R R ≥ l l ≥ j w ÷/
  
Cotter, W. (2022). The Arabic dialect of Gaza City. Journal of the International Phonetic Association, 52(1), 122-134. doi:10.1017/S0025100320000134 [Pas mis au propre]
  - Page 124 : Trill r
  - Page 126 : /r/ [r] [ˈraː.ʤɪl] ‘man’
  - Page 128 : /r/ in Gaza City Arabic is realized both as a true trilled [r], as well as a pharyngealized [r≥]. Pharyngealized [r≥] generally occurs in the environment of low vowels. Examples of pharyngealized [r≥] are readily available and can be seen in cases such as [b“r≥a] ‘outside’.
Trilled /r/ can be seen in cases such as [ra'Il] ‘man’. In connected speech contexts /r/ is also often realized as a tap [R], e.g. [nUÂRUt] ‘we farm’.

```{r}
languages_data <- languages_data %>% 
     dplyr::mutate(r_l_distinction = ifelse(Languages=="arabic",1,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="arabic",1,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="arabic",1,trill_occ))
```

## armenian

```{r}
languages_data <- languages_data %>% 
     dplyr::mutate(glottocode = ifelse(Languages=="armenian","nucl1235",glottocode))
```

Phoible :   
https://phoible.org/languages/nucl1235  
2 inventaires :  
  - SPA r allophone r   ; contrast r/l
  - UPSID r pas d'allophone  ; contrast r/l

Allen, W. Sidney. 1950. Notes on the Phonetics of an Eastern Armenian Speaker. Transactions of the Philological Society. 180–206. (Reprint by Hertford 1951).  

Wikipedia :    
https://en.wikipedia.org/wiki/Armenian_language  
  - Trill 		/r/ ռ – ṙ 	 
    Flap 		/ɾ/ ր – r 	  

Dum-Tragut, Jasmine (2009), Armenian: Modern Eastern Armenian, Amsterdam: John Benjamins Publishing Company  
  - Page 13 : ռ r̊ [r]  
              ր r [ɾ]   
  - Page 19 : Rhotics  
                    alveolar  
              flap      ɾ  
              trill     r  
In SMEA the flap n is much more wide-spread than its trill counterpart.  
The flap [ɾ] can be found in all positions, whereas the distribution of trill [r] is comparatively restricted.  
The trill [r] seldom occurs in the initial position, with the exception of some words among that ṙus [rus] “Russian” and its derivations as well as of loans with initial trill [r], [...]. In medial position, trill r [r] occurs frequently [...].  
  - Page 20 :  In final position trill r [r] is also found; also in some words which have lost their originally final n from Classical Armenian [...].  
  
Glottologue :    
https://glottolog.org/resource/languoid/id/nucl1235  

Autres sources :    

Catford, J. (2001). On Rs, rhotacism and paleophony. Journal of the International Phonetic Association, 31(2), 171-185. doi:10.1017/S0025100301002018  
  - Page 180 : The case of Armenian  
               Classical (and at least some modern dialects of ) Armenian have two r-phonemes ± a trill and a fricative or, perhaps less commonly, a single tap. The untrilled r appears to be what Armenian normally inherited from Indo-European, the trill having various secondary sources  

Seyfarth, S., Dolatian, H., Guekguezian, P., Kelly, N., & Toparlak, T. (2023). Armenian (Yerevan Eastern Armenian and Beirut Western Armenian). Journal of the International Phonetic Association, 1-34. doi:10.1017/S0025100323000130
  - Page 3 : Trill (r)
             Tap ɾ
  - Page 4 : Yerevan Armenian also has a phonemic trill, which is merged to the tap in Beirut Armenian^5
            [5] The Armenian tap has a fricative variant in syllable codas, discussed in Sonorants. We chose the tap as the basic symbol because (i) the tap has a wider distribution, (ii) a tap spirantization process seems more phonetically natural than a fricative tapping process, (iii) coauthor PG reports that some Armenian varieties do not have a fricative variant, and (iv) the tap follows the convention in Armenian linguistics literature.
  - Page 5 : r rɑmkɑˈkɑn ռամկական ‘democratic’
  - Page 6 : ɾ   ɾɑˈfi  ɾɑˈfi    Րիֆֆա    'Raffi (name')
  - Page 15 : The rhotics /r, R/ are contrastive in Yerevan Armenian, with minimal pairs such as the following:
              [examples]
               forms and in emphatic speech.Footnote 8 For example, SK’s recordings contrast /ˈseɾ/ and /ˈser/ primarily via louder frication noise for the tap /ɾ/ with a low spectral peak around 4 kHz (perhaps non-sibilant postalveolar fricative [ɹ̝̊])Footnote 9. However, word-final /ɾ/ is often unambiguously tapped or approximated in connected speech, such as in the accompanying recordings of the North Wind and Sun by HD and SK (see Transcription of recorded passage). Tap and approximant variants are both common in intervocalic position. The recordings of /kʰɑˈɾi/ ‘rock (DAT.GEN)’ and /χɑˈɾisχ/ ‘anchor’ contributed by SK illustrate approximant variants of /ɾ/.

The recording of /ˈser/ is not trilled, but has a falling third formant that is excited by aspiration without voicing, and only weak or absent oral frication noise. Most of the coda trills in the accompanying recordings contributed by SK are voiceless and/or spirantized. For example, the recording of /ˈ t͡sɑr/ ‘tree’ in the section Vowels (below) includes a voiceless trill, while the recording of the same word in the Consonants table is more similar to the variant in /ˈser/. A voiced trill coda is illustrated in the recording of /ˈkor/ ‘coerced labor (archaic)’, from the speaker who contributed /ˈkoʁ/ ‘rib’ in the subsection Yerevan Armenian plosives.

Although the two rhotics are contrastive in Yerevan Armenian, the tap is sometimes trilled before coronals (Xačatryan Reference Xačatryan1988: 108; Vaux Reference Vaux1998: 19; also noted by speaker VP), and word-initial trills are often reduced to the tap.Footnote 10

In Western Armenian, both sounds are merged to the tap /ɾ/ in all environments (Vaux Reference Vaux1998), though they are prescriptively taught as contrastive in Canadian Armenian language schools (Talia Tahtadjian, p.c.) and some Western Armenian dictionaries (Sak’apetoyean Reference Sak’apetoyean2011; also in Fairbanks Reference Fairbanks1948). For the Western Armenian community in Canada, Tahtadjian (Reference Tahtadjian2020) reports that the Western Armenian tap and trill are acoustically distinguishable: the Western trill sometimes has multiple articulator contacts (occurring in about 30 of onset trills produced by older speakers, and <15 of other trills), and the trill is about 2 ms longer in onset position and 4 ms longer in coda position. These differences are probably too small or too variable to reliably index a category difference. In the accompanying recording of Beirut Armenian [rɑzmɑˈɡɑn] ‘military’ contributed by HD, the onset consonant has multiple contacts, but HD indicates that the tap and trill were perceptually indistinguishable for his peer group in a Lebanese Armenian school.

In both dialects, the two rhotics are rare word-initially. While there are some native rhotic-initial words such as Yerevan /rɑmkɑˈkɑn/ ‘democratic’, most such words are names like Yerevan /rɑfɑˈjel/ ‘Raphael (name)’ and borrowings such as Yerevan /rɑˈbi/ ‘rabbi’. For these words, the Beirut forms use a tap instead of the trill: /ɾɑmɡɑˈɡɑn/, /ɾɑfɑˈjel/ (), /ɾɑˈpʰi/. Both rhotics are more frequent medially and finally, and occur in Yerevan minimal pairs such as the above.

Most sources treat /j/ as phonemic (in contrast with Vaux Reference Vaux1998), though it has a limited distribution. In word-initial position, the palatal approximant /j/ primarily occurs before /e/, such as in /ˈjeɾpʰ/ ‘when’ (see Vaux Reference Vaux1998: 13 for a phonological analysis and Dum-Tragut Reference Dum-Tragut2009: 14–17 for a lexical catalog; see also Vowels). All attested word-initial /jɑ/ are borrowings, such as /ˈjɑvɾəm/ , ‘my dear’ from Turkish. It is also found word-initially before the back vowels /u/ and /o/ in a handful of native words, such as /ˈjuʁ/ , ‘oil’ and /ˈjotʰə/ , ‘seven’. In some words, the Yerevan Armenian sequence /ju/ is pronounced [ʏ] in Beirut Armenian, though there is substantial variation (Khanjian Reference Khanjian2011). For example, Yerevan /ˈɡjuʁ/ ‘village’ corresponds to Beirut [ˈkʰʏʁ] ∼ [ˈkʰjʏʁ] ∼ [ˈkʰjuʁ] .

In native words, coda /j/ does not appear after /ə/, or word-finally after /i/ or /u/. It occurs in native complex codas only after /ɑ/ and /u/, such as in /ˈhɑjɾ/ ‘father’. Word-medially, /Cj/ sequences may have ambiguous onset syllabification (Margaryan Reference Margaryan1997: 55), such as in Yerevan /sen.jɑk, se.njɑk/ and Beirut /sen.jɑɡ, se.njɑɡ/ ‘room’ (see also Syllable structure). 


```{r}
languages_data <- languages_data %>% 
     dplyr::mutate(r_l_distinction = ifelse(Languages=="armenian",1,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="armenian",1,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="armenian",1,trill_occ))
```

## asl

```{r}
languages_data <- languages_data %>% 
     dplyr::mutate(glottocode = ifelse(Languages=="asl","amer1248",glottocode))
```

Phoible : NA  

Wikipedia :  
https://en.wikipedia.org/wiki/American_Sign_Language

Glottologue :    
https://glottolog.org/resource/languoid/id/amer1248    

Autres sources : NA  

```{r}
languages_data <- languages_data %>% 
     dplyr::mutate(r_l_distinction = ifelse(Languages=="asl",as.integer(NA),r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="asl",as.integer(NA),trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="asl",as.integer(NA),trill_occ))
```

## austriangerman

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="austriangerman","uppe1398",glottocode))
#We use Dialect: Upper Austria Bavarian for the glottocode
```

Phoible : NA  

Wikipedia :   
https://en.wikipedia.org/wiki/Austrian_German

Glottologue :  
https://glottolog.org/resource/languoid/id/uppe1398

Autres sources :  

Moosmüller, S., Schmid, C., & Brandstätter, J. (2015). Standard Austrian German. Journal of the International Phonetic Association, 45(3), 339-348. doi:10.1017/S0025100315000055
  - Page 340 : Trill r
              The table presents the consonant phonemes of SAG. A speaker-specific representation has to be assumed regarding the trill. The chart lists both the uvular trill and the alveolar trill. Most speakers make use of a uvular production (either trill or fricative). However, for those speakers who exclusively apply an alveolar production (either trill or approximant), /r/ has to be assumed.
  - Page 341 : /ʀ/ [ˈʁoːsɛ] Rose ‘rose’
                   [ˈleːɐʁɐ] Lehrer ‘teacher’
                   
              SAG features a wide variety of realisations of the trill. In approximately the past 40 years, the pronunciation norm has changed from an alveolar to a uvular trill. The latter is mostly pronounced as a fricative, either voiced or voiceless. Alveolar trills are still in use, mostly pronounced as an approximant. In final position and before consonants, the /ʀ/ is vocalised to either [ɐ̯ ] or [ɐ], e.g. Vater [ˈfɑːtɛɐ̯ ] or [ˈfɑːtɐ] ‘father’ or Kirche [ˈkIɐ̯ xɛ]6 ‘church’. Preceding /ʀ/, the vowel quality of [+constricted]7 vowels usually changes to [–constricted], before r-vocalisation takes place, e.g. Moor [mɔɐ̯ ]8 ‘bog’. Reduction of the sequence er to [ɐ] is only allowed in unstressed prefixes, e.g. verkaufen [fɛɐ̯ ˈkɑɔ̯ fɱ̩] or [fɐˈkɑɔ̯ fɱ̩] ‘to sell’. However, reduction to [ɐ] is not allowed in prefixes without consonantal onset, e.g. erlauben [ɛɐ̯ ˈlɑɔ̯ bm̩ ]9 ‘to allow’. Following /ɑ/, /ʀ/ is vocalised as well, however, the result of vocalisation, [ɐ̯ ], is absorbed, e.g. Parlament [pɑːlɑˈmɛnt] ‘parliament’, rar [ʀɑː] ‘scarce’. In intervocalic position, /ʀ/ is preserved. Again, in case of a preceding [+constricted] vowel, a change in vowel quality takes place and [ɐ̯ ] emerges, e.g. Lehrer [ˈlɛɐ̯ ʀɐ] ‘teacher’.
              
  - Page 345-346 : Aucun r dans la transcription, d'autres types de r

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="austriangerman",1,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="austriangerman",0,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="austriangerman",1,trill_occ))
```

## basque

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="basque","basq1248",glottocode))
```

Phoible :   
https://phoible.org/languages/basq1248
  - SPA r allophones r ʁ; et ɾ allophones ɾ   ; contrast r/l
  - UPSID r̪|r pas d'allophones; et	ɾ̪|ɾ pas d'allophones   ; contrast r̪|r/l̪|l
  - UZ r allophones r; ɾ allophones ɾ r   ; contrast r/l
  - EA 2340 r̺  pas d'allophones; et ɾ̺ pas d'allophones  ; contrast r̺/l̺
  - EA 2349 r̺  pas d'allophones; et ɾ̺ pas d'allophones    ; contrast r̺/l̺
  - EA 2583 r̺  pas d'allophones; et ɾ̺ pas d'allophones    ; contrast r̺/l̺
  - EA 2615 r̺  pas d'allophones; et ɾ̺ pas d'allophones    ; contrast r̺/l̺
  
Wikipedia :   
https://en.wikipedia.org/wiki/Basque_language
  - Rhotic[a] 	Trill 			r-, -rr-, -r
                                /r/ 			
                Tap 			-r-, -r
                                 /ɾ/ 			
  - Basque's two rhotics only contrast when between vowels, where the trill is written as -rr- and the tap as -r-. When a suffix is added to a word ending in -r, a trill is generally used, as in ederrago 'more beautiful', from eder 'beautiful' and -ago. There is a small number of words which are exceptions to this rule, with de Rijk listing the following ten common ones: zer, ezer, nor, inor, hor, paper, plater, plazer, ur, and zur.[61]
  - Unless they are recent loanwords (e.g. Ruanda "Rwanda", radar, robot ... ), words may not have initial ⟨r⟩. In older loans, initial r- took a prosthetic vowel, resulting in err- (Erroma "Rome", Errusia "Russia"), more rarely irr- (for example irratia "radio", irrisa "rice") and arr- (for example arrazional "rational").[66]

[61] de Rijk, Rudolf P. G. (2008). Standard Basque: a progressive grammar. Cambridge, MA: MIT Press. ISBN 978-0-262-04242-0. OCLC 636283146.pp. 7–8
[66] de Rijk, Rudolf P. G. (2008). Standard Basque: a progressive grammar. Cambridge, MA: MIT Press. ISBN 978-0-262-04242-0. OCLC 636283146.p.8

Glottologue :  
https://glottolog.org/resource/languoid/id/basq1248

Autres sources :  

N'diaye, Geneviéve. 1970. Structure du dialécte basque de maya. The Hague: Mouton.  
Gavel, Henri. 1929. Grammaire Basque. Bayonne: Courrier.  
Hualde, José Ignacio and Lujanbio, Oihana. 2010. Illustrations of the IPA: Goizueta Basque. Journal of the International Phonetic Association 40. 113–127. Cambridge University Press.  
  - Page 113 : Rhotic tap ɾ <-r->  
               Rhotic trill <r->,  
                          <-rr->  
  - Page 115 : /ɾ/ Initial: —– Medial: p are /p aɾe/ ‘pair’ Final: —–  
              /r/ Initial: ro  ́ eri /ro ́ e ɾi ‘drug store’ Medial: / arr ́ ai /ar ́ ai/  ‘fish’ Final: am ́ ar /am ́ ar/ ‘ten’  
  - Page 120 : Rhotics As in Spanish, the contrast between rhotic tap and trill is only made in word-internal intervocalic position. Elsewhere the distinction is neutralized. Unlike in Spanish, this neutralization in Basque is generally in favor of the trill. Neither rhotic is found in wordinitial position in the native lexicon, although the trill is found in a few recent borrowings from Spanish, e.g. r` adio ‘radio’. Notice also ro  ́ eri /ro ́ eɾi/ ‘drugstore’ < Sp. droguer ́ıa. Phonetically (also as in Spanish), rhotics are often realized without full contact, as approximants or fricatives.            
  
Hualde, José Ignacio. 1991. Basque phonology. (Theoretical linguistics.) London and New York: Routledge. xiv+210pp. (Rev. version of author's University of Southern California dissertation Includes bibliographical references (p. [199]-206) and index).  
  - Page 7 du document : Basque orthography  
                         rr = rhotic trill, []  
                         r = rhotic flap [r] between vowels. Elsewhere the distinction between the rhotics is neutralized.  
  - Page 8 du document : Phonetic symbols used in Basque transcriptions  
                         [r] alveolar rhotic flap  
                         [r̄] alveolar rhotic trill  
  - Page 10 : r  
              r̄  
              the sonorants /r,r̄/ are alveolar.  
  - Page 11 : The two rhotics segments, flap /r/ and trill /r̄/, only contrast intervocally, as in Spanish: erre [ere] 'also', erre [er̄e] 'to burn'. Elsewhere the distinction is neutralized. Unlike Spanish, in pre-or-postconsonantal position the realization is a trilled [r̄] in many Basque dialects arto [ar̄to] 'corn', andre [andr̄e] 'woman'. In the spelling, the double grapheme indicated the trill [r̄] between vowels; r is a flap [r] intervocalically.  
  - Page 12 : Word-initially, the two rhotics, [r], [r̄] do not occur in native words. Assimilated borrowings from Spanish with initial [r̄] (orthographic r-) in the lender language undergo epenthesis, generally of [e-], although there is some variation.  
  - Page 40 : In many dialects, some long vowels may be created by the optional deletion of a consonant [... r] between two identical vowels [paratu], [paatu] 'to put'. In Arbizy, on the other hand, vowel lenght is a distinctve feature.  
  - Page 54 : GERNIKA  
              As in all other dialect, the rhotics are not found word-initially. The trill [r̄] only appears word initially in unassimilated borrowings. [...] In assimilated borrowings there is epenthesis. The epenthetic vowel is [e] more consistently than in other dialects.  
  - Page 76 : Flapping  
              In the variety ofOndarroa, intervocalic /d/ can optionally be realized as [r]. In this position, there is free variation between the two allophones [ð] and [r], although the [r] realization is preferred in rapid speech. This alternation, however, is limited to certain lexical domains, including seemingly all suffixation and cliticization of /d/-initial third person conjugated verbal forms; see (119a). On the other hand, across word-boundaries, outside of cliticization, and even across the members of a compound, only [ð] and never [r] can be the surface realization of intervocalic /d/, as shown in (119b):
              (120) Flapping  
                    /d/ -> [+son] / V_V (optional)  
  - Page 84 : If, after final-vowel truncation, the final consonant is /r/, it is changed to /l/:  
              The consonantal change in (8) reflects a historical change in the opposite direction, which changed I tor intervocalically:  
              (9) l > r / V_V  
              This historical change is readily observable in the treatment of Latin loans (the examples are from Mujika (l982a)):
              In contemporary Basque, however, rule (9) not only is not productive, but does not even reflect a lexical regularity, since words with intervocalic 0] are quite common . The lexical generalization that holds is, rather, that stems ending in /VrV/, not falling into the category to which rule (2) (Lowering) applies, change their /r/ into [1] in their derivational form. This is thus a case of historical rule inversion (cf. Vennemann (1972); see also Hyman (1975: 176-178)). This rule, like the other rules discussed above, can still be applied to new words which fulfil the necessary requirements, even though its application is by no means obligatory.  
  - Page 122 : The only coronal that is never affected by Affective Palatalization is the trill [r̄]. The other rhotic, the flap [r], shows a variable behaviour; it may become (j) or [ʎ], it may be deleted, or it may be unaffected.  

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="basque",1,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="basque",1,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="basque",1,trill_occ))
```

## belarusian

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="belarusian","bela1254",glottocode))
```

Phoible : NA  

Wikipedia :  
https://en.wikipedia.org/wiki/Belarusian_language  
https://en.wikipedia.org/wiki/Belarusian_phonology  
  - Trill r  
  
Mayo, Peter (2002), "Belorussian", in Comrie, Bernard; Corbett, G. G. (eds.), The Slavonic Languages, London: Routledge, pp. 887–946, ISBN 0-415-28078-8  

  => contrast r/l̪
  
Glottologue :    
https://glottolog.org/resource/languoid/id/bela1254  

Autres sources :   

Bird, S., & Litvin, N. (2021). Belarusian. Journal of the International Phonetic Association, 51(3), 450-467. doi:10.1017/S0025100319000288  
  - Page 453 : Trill r  
  Présence de trills dans les audios joints à l'article  

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="belarusian",1,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="belarusian",1,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="belarusian",1,trill_occ))
```

## breton

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="breton","bret1244",glottocode))
```

Phoible :   
https://phoible.org/languages/bret1244
  - SPA r allophones ʁ r ɾ r̩ ɹ ; contrast r/l
  - UPSID Null ; just a l
  - EA r pas d'allophone; r̥ pas d'allophone ; contrast r/l
  
Ternes, Elmer. 1970. Grammaire Structurale du Breton de l'Ile de Groix. Heidelberg: Carl Winter.
  - Page 1 : Vibrante dental r
  - Page 5 : l - r lout "part, partie" - rout "oue
             r - s raj "raie (poisson)" - saj "robe"
  - Page 18 : r Ce phoneme comprend deux allophones principaux, à points d'articulation entièrement différents:
              1° fricative postdorso-uvulaire sonore [ʁ].
              Cet allophone se rouve dans la seule position devant /l/.
              [examples]
              2° vibrante apico-alvéolaire : [r].
              Cet allophone principal se trouve dans toutes les positions sauf devant /l/.
              En position finale (devant #), /r/ comprend un à deux battements après lesquels une sorte de détente [etc.]
  - Page 19 : Dans les positions /CrəC/ et, plus rarement, /CərC/, /ə/ tend à se perdre, surtout dans un débit rapide. On a dans ces cas un [r̩] formant syllabe, dont la force articulatoire est en général assez importante. Il comporte un à trois battements. La variation entre [rə], [ər] et [r̩]  est libre et dépend en partie de la rapidité du débit.
  [examples]
              Les remarques suivantes sont d'un caractères général. Elles concernent la force articulatoire et le nombre des vibrations de [r] dans certaines positions par rapport à certaines autres positions. Les fluctuations de prononciation sont pourtant considérables. Les indications suivantes ne sont que la moyenne des réalisations possibles.
              - En position initiale (après #), [r] est généralement plus fortement articulé (deux à trois vibrations) que dans toutes les autres positions.
              [examples]
  - Page 20 : Dans les syllabes à accent primaire ou secondaire (pg. 3.4.), [r] est plus fortement articulé (et comprend plus de vibrations) que dans des syllabes non-accentuées. P.ex. le premier [r] est plus faible que le second dans (la noation de l'accent est non-phonémique):
              u[r]gà[r]tén "une carte"
              -Plus le [r] se trouve éloigné de l'accent primaire, plus son articulation est faible (et moins il comprend de vibrations). P.ex. le [r] du premier ùpt est plus fort que celui du second (la notation de l'accent est non-phonémique):
              ə[r]gàrtén "la carte"
              ə[r]gàləxǿyl "le tapage (que font les enfants)"
              -La force articulatoire de [r] dans le groupe /rC/ dépend de la nature de C:
              Elle est relativement forte devant /p, t, k, b, d, g/ (deux à trois vibrations)
              Elle est moyenne devant /m, n, w', w/ (une à trois vibrations)
              Elle est relativement faible devant /c^v, j^v, f, s, s^v, x, v, z, z^v, j/ une à deux vibrations
              /r/ ne forme pas de groupes avec /gn, ng,w~/. Pour /rl/, v. 1° ci-dessus (p.18).
              Exemples en syllabe à accent primaire:
              [examples]
              Exemples en syllabe à accent secondaire:
              [examples]
  - Page 21 : Exemples en syllabe non-accentuée (articulation plus faible, moins de vibrations) :
              [examples]
              - La force articulatoire de [r] dans le groupe /Cr/ est indépendante de la nature de C. Elle est relativement forte (deux à trois vibrations). Elle est un peu moins forte dans le groupe /CCr/ (une à deux vibrations).
              [examples]
              Dans les positions suivantes, /r/ comprend très rarement plus d'un seul battement: [ɾ]. Par contre, ce dernier peut varier librement, surtout dans un début rapide avec une fricative alvéolaire sonore ressemblant au [ɹ] anglais, mais avec très peu de frication, formant presque une semi-voyelle  [ɹ̯].
              - Cette variation libre avec [ɹ̯] est particulièrement fréquente devant /xl/ et /xr/. Dans ces cas, [ɹ̯] peut même varier librement avec [ø], s'amalgamant avec le /x/ suivant dans une fricative glottale sonore [ɦ] :
              [examples]
  - Page 22 : En syllable non-accentué se trouvant entre deux syllabes dont la précédente porte l'accent secondaire et la suivante l'accent primaire (la noation de l'accent est non-phonémique: v. pg. 3.4):
              [examples]
              - Dans une position relativement éloignée de l'accent primaire suivant, surtout en position préconsonantique (la notation de l'accent est non-phonémique):
              [examples]
              -En position intervolcaique (rare):
              [examples]
              
Bothorel, A. 1982. Etude Phonétique et Phonologique du Breton Parlé à Argol (Finistere-Sud). Lille: Atelier National Reproduction des Thèses, Université Lille III. [pas cherché]

Iosad, Pavel. 2013. Representation and variation in substance-free phonology: A case study in Celtic. (Doctoral dissertation, Universitetet i Tromso). [pas cherché]

Wikipedia :   
https://en.wikipedia.org/wiki/Breton_language
  - Trill 			r /r/          (r /ʁ/) 
    Approximant 	central 				(r /ɹ/) 
    The pronunciation of the letter ⟨r⟩ varies nowadays: [ʁ] is used in the French-influenced standard language and, generally speaking, in the central parts of Lower Brittany (including the south of Trégor, the west of Vannetais and virtually all parts of Cornouaille) whereas [r] is the common realisation in Léon and often in the Haut-Vannetais dialect of central Morbihan (in and around the city of Vannes and the Pays de Pontivy), though in rapid speech mostly a tapped [ɾ] occurs. In the other regions of Trégor [ɾ] or even [ɹ] may be found.
    
Glottologue :  
https://glottolog.org/resource/languoid/id/bret1244  

Autres sources : NA

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="breton",1,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="breton",0,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="breton",1,trill_occ))
```

## bulgarian

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="bulgarian","bulg1262",glottocode))
```

Phoible :    
https://phoible.org/languages/bulg1262  
4 inventaires  
  - SPA r allophone r; rʲ allophone rʲ   ; contrast 	r/lˠ (allophones of lˠ -> l lˠ)
  - UPSID r pas d'allophone; rʲ pas d'allophone   ; contrast r/l̪ˠ
  - UZ r̟ allophone r̟    ;  contrast r̟/l̟
  - EA r pas d'allophone; rʲ pas d'allophone   ; contrast r/ɫ̪ (on a aussi ʎ dans l'inventaire)
  
Klagstad, H. 1958. The Phonemic System of Colloquial Standard Bulgarian. Slavic and East European Journal 16. 42–54.  
Aronson, H. I. 1968. Bulgarian Inflectional Morphophonology. The Hague: Mouton.  
Bidwell, C. 1968. The Stress Patterns of the Noun in Bulgarian. Studies in Linguistics 20. 41–47.  
Scatton, Ernest A. 1984. A Reference Grammar of Modern Bulgarian. Columbus, Ohio: Slavica Publishers.  
Бояджиев, Тодор Апостолов and Куцаров, Иван and Йордан, Пенчев. 1999. Съвременен български език: фонетика, лексикология, словообразуване, морфология, синтаксис: учебник за студенти от филологическите факултети и педагогическите институти. Петър Берон.  


Ternes, Elmar and Vladimirova-Buhtz, Tatjana. 1990. Illustrations of the IPA: Bulgarian. Journal of the International Phonetic Association 20. 45–47. Cambridge University Press.  
  - Page 45 : Trill r  
              r roza 'rose'  
  - Page 46 : An alternative analysis postulates the following palatalized consonants: /... rʲ .../  
              /r/ is a trill.  

Wikipedia :      
https://en.wikipedia.org/wiki/Bulgarian_language  
https://en.wikipedia.org/wiki/Bulgarian_phonology  
  - Trill r	rʲ 	  

  => contrast r/l̪
  
Glottologue :     
https://glottolog.org/resource/languoid/id/bulg1262    

Autres sources : NA

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="bulgarian",1,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="bulgarian",1,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="bulgarian",1,trill_occ))
```

## calabrese

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="calabrese","sout2616",glottocode))
# Mais comme dialect of Sicilian, on recherche infos pour Sicilian
```

Phoible : NA

Wikipedia :   
https://en.wikipedia.org/wiki/Sicilian_language
  - Trill 		r 			
    Flap 		ɾ
    Approximant 		l
    
  - RR — The digraph -rr-, depending on the variety of Sicilian, can be a long trill [rː] (hereafter transcribed without the length mark)[89] or a voiced retroflex sibilant [ʐː].[86] This innovation is also found under slightly different circumstances in Polish, where it is spelled -rz-, and in some Northern Norwegian dialects, where speakers vary between [ʐ] and [ɹ̝].[citation needed] At the beginning of a word, the single letter r is similarly always pronounced double,[clarification needed] though this is not indicated orthographically. This phenomenon, however, does not include words that start with a single r resulting from rhotacism or apheresis (see below), which should not be indicated orthographically to avoid confusion with regular double r.[citation needed]
  - Rhotacism and apheresis — This transformation is characterized by the substitution of single d by r. In Sicilian this is produced by a single flap of the tongue against the upper alveolar ridge [ɾ]. This phenomenon is known as rhotacism, that is, the substitution of r for another consonant; it is commonly found both in Eastern and Western Sicilian, and elsewhere in Southern Italy, especially in Neapolitan. It can occur internally, or it can affect initial d, in which case it should not be represented orthographically to avoid confusion with the regular r (see above). Examples : pedi ("foot") is pronounced [ˈpɛːɾɪ]; Madonna ("Virgin Mary") is pronounced [maˈɾɔnna]; lu diri ("to say it") is pronounced [lʊ ˈɾiːɾɪ].[93] Similarly, apheresis of some clusters may occur in certain dialects, producing instances such as 'ranni [ˈɾannɪ] for granni "big".
  
    => contrast r/l

PROPOSTA DI NORMALIZZAZIONE ORTOGRAFICA COMUNE DELLA LINGUA SICILIANA per le varietà parlate sull’isola di Sicilia, arcipelaghi ed isole satelliti, e nell’area di Reggio Calabria (2017) - 
  - page 10 : r [ɾ]    rr [ɾ:]
  - page 18 :  Facciamo l’esempio della parola [‘ɾ:ɔta], che pur presentando in ogni caso al suo inizio il fonema [ɾ:] tenderemo sempre a rappresentare come rota (con un solo grafema r)
  - Page 25 : 8) Rappresentiamo il fonema [l] servendoci del grafema l e la sua controparte geminata /l:/
come ll. Nella stragrande maggioranza dei casi si sconsiglia l'uso del grafema l raddoppiato ad
inizio di parola, che può però essere soggetto a raddoppiamento all'interno di parola,
tranquillamente rappresentabile con -ll-;
  - Page 27 : 13) Rappresentiamo il fonema /ɾ/ attraverso il grafema r e la sua controparte geminata /ɾ:/ come
rr. Nella stragrande maggioranza dei casi si sconsiglia l'uso del grafema r raddoppiato ad inizio
di parola, che può però essere soggetto a raddoppiamento all'interno di parola, tranquillamente
rappresentabile con rr;
Esempio 1: la parola rota102 è un esempio di parola che vede all'inizio il grafema r non
geminato.
«N’ ‘o sacciu. Si non mi ritiru non ti pigghiari pena e non stari in pinseri, ca nni me
frati, sugnu!»103
(N. Martoglio, San Giuvanni decullatu (Atto I, scena X), in Teatro – Volume II, Messina,
Casa Editrice G. D’Anna, 1965)
Esempio 2: la parola arruccatu104 è un esempio di parola che vede il suono [ɾ:]
tranquillamente rappresentato dal digramma -rr-.
«Ddocu la menti mia ristà cunfusa! / dissi a ddu versu: É megliu ca ti chianti… / ‘Na vuci
arrispunnì: - Sugnu la Musa, / ti suggerisciu iu, pirchì ti scanti?»105
(G. Pace ‘Peppipaci’, Mascari di Paci, Canicattì, Tipografia Modena, 1937)

Glottologue :  
https://glottolog.org/resource/languoid/id/sici1248
https://glottolog.org/resource/languoid/id/sout2616

Joseph F. Privitera. 1998. Basic Sicilian: A brief reference grammar. Lewiston: Edwin Mellen. x+67pp.
  - Page viii : b, f, l, n, q, t, v are prononced as in English
                r, when single, is formed with a quick flip of the tongue on the palate behind the front teeth: as in caru, dear; puru, pure.
                  When r is double, or when it is the first letter in a word, the tongue is suspended between the roof and the bottom of the mouth, not touching either, its sides are curved agaisnt the sides of the mouth, whil the tip is rounded to let the air escape: as in rota, wheel; carru, cart
                

Autres sources :  

Antonio Pisano, Dizionario dialettale calabrese, 2011
  - Traduit avec GoogleTrad en ligne :
  - Page 3 du doc : Erre – R – double la consonne suivante pour la logique de notre prononciation dure, forte et marquée : scarro > scarppàru.

  => on a une entrée pour le r et pour le l; le rr est présent dans le dictionnaire
  
https://soundcomparisons.com/#/en/Romance/languagesXwords/Lgs_Sln/Wds_Sln
On check sur SoundComparisons:
  - Calabria Lausb.: Oriolo [ˈʰɾaˑpə̆ðə̤] [ˌpɾɪ̆st̪ɘ̆ˈ] [ˈtɛːɹ̝ɹ̝ə] [ˈʰɾɔotʰə]
  - Calabria N.: Cetraro [ˈzvɛ̞ˑlt̪ʰo̞] [ˈtɛ̞ˑrrɐ] [ˈrɔ̞ːta]
  - Calabria S.: Cropani [ˈlˠestʰʊ̆] [ʔˈlˠestʰʊ̆] [ˈt̪ʰɛˑrrɐ] [ˈrɔːta̝]

Le trill existe pour la géminée ?

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="calabrese",1,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="calabrese",0,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="calabrese",1,trill_occ))
```

## catalan

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="catalan","stan1289",glottocode))
```

Phoible :     
https://phoible.org/languages/stan1289  
4 inventaires  
  - PH r allophone r; ɾ allophone ɾ   ; contrast r/l
  - EA2425 r̺ pas d'allophone; ɾ̺ pas d'allophone ; contrast r̺/l̺
  - EA2555 r̺ pas d'allophone; ɾ̺ pas d'allophone ; contrast r̺/ɫ̺
  - EA2594 r̺ pas d'allophone; ɾ̺ pas d'allophone ; contrast r̺/ɫ̺

Carbonell, Joan F. and Llisterri, Joaquim. 1992. Illustrations of the IPA: Catalan. Journal of the International Phonetic Association 22. 53–56. Cambridge University Press.  
  - Page 53 : Trill r  
              Tap or Flap ɾ  
              
              r serra 'saw' (n.)  
              ɾ cera 'wax'  
Lloret, Maria-Rosa. 2011. La fonologia del catal`a. Santillana, Grup Promotor.  

Wikipedia :    
https://en.wikipedia.org/wiki/Catalan_language  
  - Tap   ɾ  
  - Trill r  
  - The distribution of the two rhotics /r/ and /ɾ/ closely parallels that of Spanish. Between vowels, the two contrast, but they are otherwise in complementary distribution: in the onset of the first syllable in a word, [r] appears unless preceded by a consonant. Dialects vary in regards to rhotics in the coda with Western Catalan generally featuring [ɾ] and Central Catalan dialects featuring a weakly trilled [r] unless it precedes a vowel-initial word in the same prosodic unit, in which case [ɾ] appears.[104]  

Padgett, Jaye (December 2009). "Systemic Contrast and Catalan Rhotics". The Linguistic Review. 26 (4): 431–. doi:10.1515/tlir.2009.016. S2CID 15197551  

Glottologue :    
https://glottolog.org/resource/languoid/id/stan1289  

Autres sources :  NA

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="catalan",1,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="catalan",1,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="catalan",1,trill_occ))
```

## chinese

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="chinese","mand1415",glottocode))
#By Chinese I'll assume Mandarin Chinese
```

Phoible :   
https://phoible.org/languages/mand1415
4 inventories :
  - SPA i	allophones	i̥ ɻ̩ i s̞̩ ɻ̥̩ z̞̩ ; ʐ allophones ʐ;    presence du l allophone l
  - UPSID ʐ pas d'allophone ; presence du l 
  - PH ɹ allophone ɹ ;  presence du l allophone l
  - EA ɹ̺ pas d'allophone ; ɹ̪̩ pas d'allophone  ; presence du l

Chao, Yuan-Ren. 1968. A Grammar of Spoken Chinese. Berkeley: University of California Press.
Karlgren, Berhard. 1926. Etudes sur la phonologie chinoise. (Archives d'Etudes Orientales.) K. W. Appelberg/E. J. Brill.
Cheng, C.-C. 1973. A Synchronic Phonology of Mandarin Chinese. The Hague: Mouton.
Dow, Francis D. M. 1972. An Outline of Mandarin Phonetics. Canberra: Faculty of Asian Studies, Australian National University.
Hanyu, Fangyan Gaiyao. 1960. Outline of Chinese Dialects. Beijing: Wenzi Gaige Chubanshe.
Jiahua, Y. 1960. Hanyu Fangyan Gaiyao [An Outline of the Chinese Dialects]. Beijing: Wenzi Gaige Chubanshe.


Lee, Wai-Sum and Zee, Eric. 2003. Illustrations of the IPA: Standard Chinese (Beijing). Journal of the International Phonetic Association 33. 109–112. Cambridge University Press.
  - Page 109 : Approximant ɹ̺
              ɹ̺ ɹ̺ən person
  - Page 110 : Syllabic consonants
              ɹ̩ sɹ̩ to think
                ʃɹ̩ poem
  
Wikipedia :   
https://en.wikipedia.org/wiki/Mandarin_Chinese
  - Sonorants 	/ɻ/ ~ /ʐ/ ⟨r⟩ 
  - In some dialects, initial /l/ and /n/ are not distinguished. In Southwestern Mandarin, these sounds usually merge to /n/; in Lower Yangtze Mandarin,  they usually merge to /l/.[94]
People in many Mandarin-speaking areas may use different initial sounds where Beijing uses initial r- /ɻ/. Common variants include /j/, /l/, /n/ and /w/.[93]
  - Syllabic consonants

The syllables written in pinyin as zi, ci, si, zhi, chi, shi, ri may be described as a sibilant consonant (z, c, s, zh, ch, sh, r in pinyin) followed by a syllabic consonant (also known as apical vowel in classic literature):
    [ɹ̩ ~ z̩], a laminal denti-alveolar voiced continuant,[a] in zi, ci, si ([tsɹ̩ tsʰɹ̩ sɹ̩]);
    [ɻ̩ ~ ʐ̩], an apical retroflex voiced continuant,[a] in zhi, chi, shi, ri ([ʈʂɻ̩ ʈʂʰɻ̩ ʂɻ̩ ɻɻ̩]).

https://en.wikipedia.org/wiki/Standard_Chinese_phonology
  - Liquid 		l 	(ɻ) 
  - /l/ 	Like English clear l, as in RP lay (never dark, i.e. velarized)
  - /ʐ/ ([ʐ ~ ɻ])[a] 	Similar to z in zoo in English, but with a retroflex articulation. L2 learners may pronounce it as an English R, but lips are unrounded. 
   For pronunciation in syllable-final position, see § Rhotic coda. 
Glottologue :  
https://glottolog.org/resource/languoid/id/mand1415

Autres sources :  

https://www.youtube.com/watch?v=PQByGg0uAC4
Los CHINOS SÍ pueden PRONUNCIAR la ERRE  -> City : Danyang https://en.wikipedia.org/wiki/Dangyang
-> So once I found this blog:
http://www.lawview.net/Item/Show.asp?m=1&d=2009
The mystery of the popping tongue in the Hubei Jing (Men) and Dang (Yang) dialects
Author: Anonymous Editor: 本站 Published: Chen Huilin Published: 2021-02-24 10:06 Last
modified: 2021-02-24 10:06 Read: 911 times
[I've translated with GoogleTrad to get the overall meaning]
  - 3 endroits pour le trill en Chine : firstly, in the Xuanzhou small area of the Wu dialect region, secondly, in the Hubei small area and
part of the Hubei small area of the Southwest official dialect region, and thirdly, in part of the Yu southeast small area of the Central Plains official dialect region.
  - The presence of the popping tongue in the dialects of Jingdang can only be the result of the influence of languages outside the Sino-Tibetan family.

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="chinese",0,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="chinese",0,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="chinese",1,trill_occ))
```

## creole

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="creole",as.character(NA),glottocode))
```

Phoible : --

Wikipedia : --   

Glottologue : --

Autres sources : --

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="creole",as.character(NA),r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="creole",as.character(NA),trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="creole",as.character(NA),trill_occ))
```

## croatian

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="croatian","croa1245",glottocode))
```

Phoible :   
https://phoible.org/inventories/view/1139
  - r allophone r ; contrast r/l

Dankovičová, Jana. 1997. Illustrations of the IPA: Czech. Journal of the International Phonetic Association 27. 77–80. Cambridge University Press. [pas mis au propre]
  - Page 77 : Trill  r r̝
  - Page 78 : There are two trills, Irl, as in ruka /ruka/ 'hand', and Irl, as in reka /reka/ 'river'. The first one is an alveolar apical trill with 1-3 periods of vibration. It is immune from voicing assimilation and occurs voiced in all positions in the word. In case of Irl, the place of articulation is normally similar to that for Irl. Although it may be produced with the blade of the tongue, the main differentiating characteristic from Irl is the number of vibrations, which may be 1-2 greater than in Irl, and their lesser amplitude than for the vibrations in Irl. Also, the constriction is narrower and the velocity of air greater. This sound often starts as a trill but continues as a fricative and thus probably the best term for it is 'alveolar trill
fricative' with the symbol Irl. (The laminal diacritic Irl used in Ladefoged & Maddieson (1996) does not capture the sound's defining property). A voiceless allophone of Irl occurs in places where voicing assimilation applies.

Wikipedia :   
https://en.wikipedia.org/wiki/Croatian_language
https://en.wikipedia.org/wiki/Serbo-Croatian_phonology
  - Trill 		r
  - /r/ can be syllabic, short or long, and carry rising or falling tone, e.g. kȓv ('blood'), sȑce ('heart'), sŕna ('deer'), mȉlosr̄đe ('compassion'). It is typically realized by inserting a preceding or (more rarely) succeeding non-phonemic vocalic glide.
  
Glottologue :  
https://glottolog.org/resource/languoid/id/croa1245

Autres sources : NA 

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="croatian",1,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="croatian",1,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="croatian",1,trill_occ))
```

## czech

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="czech","czec1258",glottocode))
```

Phoible :    
https://phoible.org/languages/czec1258  
3 inventaires  
  - PH 	r̝ allophone 	r̝; r allophone r r̩   ; contrast r/l
  - EA2274 r pas d'allophone; r̝ pas d'allophone   ; contrast r/l
  - EA2480 r pas d'allophone; r̝ pas d'allophone   ; contrast r/l

Dankovičová, Jana. 1997. Illustrations of the IPA: Czech. Journal of the International Phonetic Association 27. 77–80. Cambridge University Press.  
  - Page 77 : Trill r  
                    r̝  

There are two trills, /r/, as in ruka /ruka/ 'hand', and /r̝/, as in řeka /r̝eka/ 'river'. The first one is an alveolar apical trill with 1-3 periods of vibration. It is immune from voicing assimilation and occurs voiced in all positions in the word. In case of /r̝/, the place of articulation is normally similar to that for /r/. Although it may be produced with the blade of the tongue, the main differentiating characteristic from /r/ is the number of vibrations, which may be 1-2 greater than in /r/, and their lesser amplitude than for the vibrations in Irl. Also, the constriction is narrower and the velocity of air greater. This sound often starts as a trill but continues as a fricative and thus probably the best term for it is 'alveolar trill fricative' with the symbol /r̝/. (The laminal diacritic /r̻/ used in Ladefoged & Maddieson (1996) does not capture the sound's defining property). A voiceless allophone of /r̝/ occurs in places where voicing assimilation applies.  

Šimáčková, Šárka and Jonáš Podlipský, Václav. 2012. Illustrations of the IPA: Czech spoken in Bohemia and Moravia. Journal of the International Phonetic Association 42. 225–232. Cambridge University Press.  
  - Page 226 : Trill r   
                     r̝  
  - Page 227 : Intervocalically, /d/ is sometimes reduced to [R] (Machaˇc & Skarnitzl 2009: 37–38)  
               Both /r/ and /r̝/ are trills though commonly realized with a single contact. Phonetically, the sound /r̝/ is a period of friction interrupted at the beginning by a contact or contacts created by a retracted apico-alveolar gesture (see Figure 1).  
               Sonorants /r/, /l/, and marginally also /m/ and /n/, become syllabic between two consonants or after a consonant at the end of a word  

Wikipedia :    
https://en.wikipedia.org/wiki/Czech_language  
  - Trill 	plain 		r ⟨r⟩  
  - Trill   fricative 		r̝ ⟨ř⟩  
  - The phoneme represented by the letter ř (capital Ř) is very rare among languages and often claimed to be unique to Czech, though it also occurs in some dialects of Kashubian, and formerly occurred in Polish.[39] It represents the raised alveolar non-sonorant trill (IPA: [r̝]), a sound somewhere between Czech r and ž (example: "řeka" (river)   (help·info)),[40] and is present in Dvořák. In unvoiced environments, /r̝/ is realized as its voiceless allophone [r̝̊], a sound somewhere between Czech r and š.[41]  
  - The consonants /r/, /l/, and /m/ can be syllabic, acting as syllable nuclei in place of a vowel. Strč prst skrz krk ("Stick [your] finger through [your] throat") is a well-known Czech tongue twister using syllabic consonants but no vowels.[42]   

Glottologue :    
https://glottolog.org/resource/languoid/id/czec1258  

Autres sources : NA  

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="czech",1,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="czech",1,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="czech",1,trill_occ))
```

## danish

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="danish","dani1285",glottocode))
```

Phoible :    
https://phoible.org/languages/dani1285  
2 inventaires  
  - UZ r (marginal) allophones r ɐ̯ ʌ̯;	ʁ̞ allophones	ʁ̞    ;  contrast r/l
  - EA ʁ pas d'allophone  ;  contrast  ʁ/l
  
Grønnum, Nina. 1998. Illustrations of the IPA: Danish. Journal of the International Phonetic Association 28. 99–105. Cambridge University Press.  
  - Page 99 : Approximant ʁ  
  - Page 100 : Add to this the non-syllabic manifestation of /r/ in post-vocalic position, transcribed either [ɐ̯] or (as in this paper) [ʌ̯].  
               Note that in a narrower transcription [...] since they are in fact lenis. [...] And [ð] and [ʁ] would be [ð̞] and [ʁ̞], respectively, for the same reason.  
               
Wikipedia :     
https://en.wikipedia.org/wiki/Danish_language  
  - Approximant  ʁ  
  - In onset /r/ is realized as a uvu-pharyngeal approximant, [ʁ], but in coda it is either realized as a non-syllabic low central vowel, [ɐ̯] or simply coalesces with the preceding vowel. The phenomenon is comparable to the r in German or in non-rhotic pronunciations of English. The Danish realization of /r/ as guttural – the so-called skarre-r – distinguishes the language from those varieties of Norwegian and Swedish that use trilled [r]. Only very few, middle-aged or elderly, speakers of Jutlandic retain a frontal /r/ which is then usually realised as a flapped [ɾ] or approximant [ɹ].   

https://en.wikipedia.org/wiki/Danish_phonology  
  - Syllable-initially, /r/ is a voiced uvular fricative [ʁ] or, more commonly, an approximant [ʁ̞].[25] According to Nina Grønnum, the fricative variant is voiceless [χ].[26] Its precise place of articulation has been described as pharyngeal,[27] or more broadly, as "supra-pharyngeal".[28] When emphasizing a word, word-initial /r/ may be realized as a voiced uvular fricative trill [ʀ̝].[16] In syllable-final position, /r/ is realized as [ɐ̯].[2]  
The alveolar realization [r] of /r/ is very rare. According to Torp (2001), it occurs in some varieties of Jutlandic dialect, and only for some speakers (mostly the elderly). The alveolar realization is considered non-standard, even in classical opera singing – it is probably the only European language in which this is the case.[29] According to Basbøll (2005), it occurs (or used to occur until recently) in very old forms of certain conservative dialects in Northern Jutland and Bornholm.[30]   

Glottologue :    
https://glottolog.org/resource/languoid/id/dani1285  

Autres sources : NA  
```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="danish",1,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="danish",0,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="danish",1,trill_occ))
```

## dutch

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="dutch","dutc1256",glottocode))
```

Phoible :    
8 inventaires :  
https://phoible.org/languages/dutc1256
  - PH ʀ allophones ʀ r   ; contrast ʀ/l
  - UZ2169 r	allophones	r r̥ ʀ ʀ̥ ʁ̞   ; contrast r/l
  - UZ2170 r	allophones	r ʀ ɾ ɢ̆     ; contrast r/l
  - UZ2171 ʀʁ	allophones	ʀʁ ʀ̥ʁ̥   ; contrast ʀʁ/l
  - UZ2172 ʀʁ	allophones	ʀʁ ʀ̥ʁ̥    ; contrast ʀʁ/l
  - UZ2173	ʀ	allophones	ʀ χ    ; contrast ʀʁ/l
  - UZ2174 	r	allophones	r ɾ r̞ ɹ̞ ɾɹ̞    ; contrast r/l
  - EA 	ɾ pas d'allophone     ; contrast ɾ/l
  
Verhoeven, Jo. 2007. Illustrations of the IPA: The Belgian Limburg dialect of Hamont. Journal of the International Phonetic Association 37. 219–225. Cambridge University Press.  
  - Page 219 : Trill R  
  - Page 220 : The voiced trill is typically uvular with a clear transitional aspect of articulation (Laver 1994). The devoiced uvular trill in word-final position is realised as a uvular fricative either with or without a transitional aspect.  

Verhoeven, Jo. 2005. Illustrations of the IPA: Belgian Standard Dutch. Journal of the International Phonetic Association 35. 243–247. Cambridge University Press.  
  - Page 243 : Trill (r)  R  
  - Page 245 : Trills  
The phonological system of Belgian Dutch has free variation between an alveolar and uvular trill. The alveolar trill is most frequent and geographically most widely distributed. The uvular trill is regionally confined to the cities of Ghent and Brussels and the province of Limburg, but estimates suggest that it is gaining fast in popularity (Van Reenen 1994). The alveolar/uvular trill is voiced in all positions, except word-finally as in [do:R9] ‘through’ and when followed by a voiceless stop in consonant clusters as in [va:R9t] ‘sails’. The phonetic characteristics of the uvular trill in the province of Limburg were investigated in Verhoeven (1994). It was found that uvular trills are virtually always realised with a clear transitional aspect of articulation, i.e. the uvula does actually trill against the back of the tongue. Only in a very small number of cases the uvular trill was realised phonetically as a voiced uvular approximant or a voiceless uvular fricative, i.e. without a transitional aspect. This limited phonetic variability of r-realisation in Belgian Dutch contrasts with the Netherlands where /r/ is also frequently realised as amongst others a central vowel, a retroflex approximant, an alveolar tap, an alveolar approximant or a uvular fricative (Van de Velde 1994).  
  
Peters, Jörg. 2006. Illustrations of the IPA: The dialect of Hasselt. Journal of the International Phonetic Association 36. 117–124. Cambridge University Press.  
  - Page 117 : Trill r  
  - Page 118 : The dialect has free variation between an alveolar and a uvular trill.  
  
Gussenhoven, Carlos and Aarts, Flor. 1999. Illustrations of the IPA: The dialect of Maastricht. Journal of the International Phonetic Association 29. 155–166. Cambridge University Press.  
  - Page 155 : Trill  R  
  - Page 156 : /R/ is a (pre-)uvular trill with a fricative component, the latter element being particularly prominent in the coda, where the consonant is partially devoiced.  

Heijmans, Linda and Gussenhoven, Garlos. 1998. Illustrations of the IPA: The Dutch dialect of Weert. Journal of the International Phonetic Association 28. 107–112. Cambridge University Press.  
  - Page 107 : Trill  R  
  - Page 108 : [ʀ] is a uvular or pre-uvular trill with a fricative component, the latter element being particularly prominent in the coda, where the consonant is partially devoiced.  
  
Peters, Jörg. 2010. Illustrations of the IPA: The Flemish-Brabant dialect of Orsmaal-Gussenhoven. Journal of the International Phonetic Association 40. 239–246. Cambridge University Press.  
  - Page 239 : Trill r  
  - page 240 : /r/ before a stressed vowel in word-initial syllables is an apicoalveolar trill or fricative. Intervocalic /r/ and /r/ in the onset after a consonant may be reduced to [ɾ]. Word-final /r/ is highly variable both within and between speakers. The most frequent variants are the apicoalveolar fricative trill [r6], the apicoalveolar fricative [®6], and an apicoalveolar affricate [ɾ®6]; the last two variants tend to become voiceless in pre-pausal position (cf. Whitley 2003).  
  
Wikipedia :     
https://en.wikipedia.org/wiki/Dutch_language#Phonology  
  - Rhotic 		r  
  - The realization of /r/ phoneme varies considerably from dialect to dialect and even between speakers in the same dialect area. Common realisations are an alveolar trill [r], alveolar tap [ɾ], uvular trill [ʀ], voiced uvular fricative [ʁ], and alveolar approximant [ɹ].  

https://en.wikipedia.org/wiki/Dutch_phonology  
  - Rhotic 		r  
  - The realization of /r/ phoneme varies considerably from dialect to dialect and even between speakers in the same dialect area:  
    The historically original pronunciation is an alveolar trill [r], with the alveolar flap [ɾ] as a common allophone.  
    The uvular trill [ʀ] is a common alternative, found particularly in the central and southern dialect areas. Uvular pronunciations appear to be gaining ground in the Randstad.[16] Syllable-finally, it may be vocalized to [ɐ], much as in German. This is more common in the (south)eastern areas (Limburg, southeast Brabantian, Overijssel).
    The coastal dialects of South Holland produce a voiced uvular fricative [ʁ].  
    The velar bunched approximant [ɹ̈] (the Gooise R, which sounds similar to the retroflex approximant) is found at the end of a syllable in the pronunciation of some speakers in the Netherlands, especially those from the Randstad, but not in Belgium. Its use has been increasing in recent years.[17]  
    
    [16] Collins, Beverley; Mees, Inger M. (2003) [First published 1981]. The Phonetics of English and Dutch (5th ed.). Leiden: Brill Publishers. ISBN 9004103406.  
    [17] Sebregts, Koen (2014). The Sociophonetics and Phonology of Dutch r (PDF) (dissertation). Utrecht: LOT. ISBN 978-94-6093-161-1. Archived (PDF) from the original on 4 March 2016. Retrieved 25 March 2017.  
    
Glottologue :    
https://glottolog.org/resource/languoid/id/dutc1256  

Autres sources : NA  

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="dutch",1,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="dutch",1,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="dutch",1,trill_occ))
```

## english

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="english","stan1293",glottocode))
```

Phoible :   
https://phoible.org/languages/stan1293
9 inventories :
  - SPA ɻ allophones 	ʐ͇ ɾ ɻ̥ ɻ ; contrast ɻ/l
  - UZ2175 ɹ allophone ɹ ; 	tʰ allophone tʰ t ɾ tʰɪ̯  ; n allophones n ɾ nɪ̯  ;	d allophones d d̥ ɾ dɪ̯ ;  contrast ɹ/l
  - UZ2176 ɹ allophone ɹ ; 	tʰ allophone tʰ t ɾ tʰɪ̯  ; n allophones n ɾ nɪ̯  ;	d allophones d d̥ ɾ dɪ̯ ;  contrast ɹ/l
  - UZ2177 ɹ allophones ɹ ɹ̥ ; 	d allophones d ɾ dʔ dⁿ dˡ d̥ ; 	tʰ allophones tʰ t t̚ ɾ tʔ ʔ tⁿ tˡ ;   contrast ɹ/l
  - UZ2178 ɹ allophones ɹ ɾ  ; tʰ allophones  tʰ t ts θ ð h ʰsʰ ɹ ɾ ʔ   ; contrast ɹ/l
  - UZ2179 ɹ allophones 	ɹ ɾ z s ʋ  ; ts allophones ts tʰ ɾ t tʔ ʔt ʔ ; contrast ɹ/l
  - UZ2180 ɹ allophones ɹ ʋ w ʔ  ; contrast  ɹ/l
  - EA2252 ɹ pas d'allophones ; contrast ɹ/l
  - EA2515 r pas d'allophones ; contrast r/l

Trnka, Bohumil. 1968. A Phonological Analysis of Present-day Standard English. (Alabama Linguistic and Philological Series, 17.) Alabama: University of Alabama Press.
Halle, Morris. 1973. Stress Rules in English: A New Version. Linguistic inquiry 4. 451–464.
Fudge, Erik. 1975. English Word Stress: An Examination of Some Basic Assumptions. In Goyvaerts, Didier L. and Pullum, Geoffrey K. (eds.), Essays on the Sound Pattern of English, 277–323. Ghent: E. Story-Scientia P.V.B.A.
O'Conner, J. D. 1973. Phonetics. Middlesex: Penguin.
Gimson, A. C. 1962. An Introduction to the Pronunciation of English. London: Edward Arnold.
Jones, D. 2003. English Pronouncing Dictionary. (16.) In Peter Roach, James Hartman and Setter, Jane (eds.) Cambridge University Press.

[I won't check the following articles because I've already checked them in the past.]

Ladefoged, Peter. 1989. Illustrations of the IPA: American English. Journal of the International Phonetic Association 19. 41–44. Cambridge University Press.

Hillenbrand, James M. 2003. Illustrations of the IPA: American English: Southern Michigan. Journal of the International Phonetic Association 33. 121–126. Cambridge University Press.

Cox, Felicity and Palethorpe, Sallyanne. 2007. Illustrations of the IPA: Australian English. Journal of the International Phonetic Association 37. 341–350. Cambridge University Press.

Watson, Kevin. 2007. Illustrations of the IPA: Liverpool English. Journal of the International Phonetic Association 37. 351–360. Cambridge University Press.

Bauer, Laurie and Warren, Paul. 2007. Illustrations of the IPA: New Zealand English. Journal of the International Phonetic Association 37. 97–102. Cambridge University Press.

Watt, Dominic and Allen, William. 2003. Illustrations of the IPA: Tyneside English. Journal of the International Phonetic Association 33. 267–271. Cambridge University Press.

Wikipedia :   
https://en.wikipedia.org/wiki/English_language
  - Approximant 						ɹ* 
    * Conventionally transcribed /r/
https://en.wikipedia.org/wiki/English_phonology
  - Approximant 	r
    This phoneme is conventionally transcribed with the basic Latin letter ⟨r⟩ (the IPA symbol for the alveolar trill), even though its pronunciation is usually a postalveolar approximant [ɹ̠]. The trill exists but is rare, found only in some Scottish, Welsh,[8] South African[9] and Indian[10] dialects. See Pronunciation of English /r/.

https://en.wikipedia.org/wiki/Pronunciation_of_English_/r/
  - "Trilled" or "Rolled" R: alveolar trill [r] (occurs in some very conservative Scottish English, South African English, some Welsh English,[4] Indian English[3] and Jersey English)
    
Glottologue :  
https://glottolog.org/resource/languoid/id/stan1293

Autres sources : NA  

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="english",1,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="english",0,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="english",1,trill_occ))
```

## esperanto

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="esperanto","espe1235",glottocode))

#J'ai pas approfondi dans les détails de la langue
```

Phoible : NA  

Wikipedia :   
https://en.wikipedia.org/wiki/Esperanto
  - Trill 	  	r 	
  - The sound /r/ is usually rendered as an alveolar trill [r], but can also be a uvular trill [ʀ],[80] a uvular fricative [ʁ],[81] and an alveolar approximant [ɹ].[82] Many other forms such as an alveolar tap [ɾ] are done and accepted in practice.
  
  [80] "PMEG – Bazaj elparolaj reguloj – Konsonanta variado". PMEG. Archived from the original on February 13, 2015. Retrieved January 31, 2018.
  [81] "Fundamento de Esperanto – Gramatiko Franca". Akademio de Esperanto. Archived from the original on February 1, 2018. Retrieved January 31, 2018.
  [82] "Fundamento de Esperanto – Gramatiko Angla". Akademio de Esperanto. Archived from the original on May 30, 2019. Retrieved January 31, 2018.
   
Glottologue :  
https://glottolog.org/resource/languoid/id/espe1235

Autres sources : NA

```{r}
languages_data <- languages_data %>%
     dplyr::mutate(r_l_distinction = ifelse(Languages=="esperanto",1,r_l_distinction))
 
languages_data <- languages_data %>%
     dplyr::mutate(trill_real = ifelse(Languages=="esperanto",1,trill_real))
 
languages_data <- languages_data %>%
     dplyr::mutate(trill_occ = ifelse(Languages=="esperanto",1,trill_occ))
```

## estonian

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="estonian","esto1258",glottocode))
```

Phoible :   
https://phoible.org/languages/esto1258
2 inventaires :
  - UZ 	r allophones r r̥ ; rː allophones rː ; 	rˑ allophones 	rˑ ; contrast r/l
  - EA 	r pas d'allophone ; contrast r/l


Asu, Eva Liina and Teras, Pire. 2009. Illustrations of the IPA: Estonian. Journal of the International Phonetic Association 39. 367–372. Cambridge University Press.
  - Page 367 : Trill r
               r raba [rapa] ‘bog’
  - Page 368 : 
  Voiced consonants [m n r v l] become unvoiced in word-final position after [t], [h] or [s]
  All word-internal and word-final alveolar consonants (except /r/) can be palatalized.
  
Hebedová, Petra. 2013. Struvcná gramatika estonvstiny. Masarykova univerzita.

Wikipedia :   
https://en.wikipedia.org/wiki/Estonian_language
  - Trill 		r

Glottologue :  
https://glottolog.org/resource/languoid/id/esto1258

Autres sources : NA  

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="estonian",1,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="estonian",1,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="estonian",1,trill_occ))
```

## farsi

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="farsi","west2369",glottocode))
```

Phoible :   
https://phoible.org/languages/west2369  
4 inventaires :  
  - SPA r allophones ɾ r r̥      ; contrast r/l
  - USPID r pas d'allophone    ; contrast r/l
  - UZ r	allophones r ɾ ɹ    ; contrast r/l
  - EA r̪ pas d'allophone   ; contrast r̪/l̪

Obolensky, Serge and Panah, Kambiz Yazdan and Nouri, Fereidoun Khaje. 1963. Persian Basic Course. Washington, D.C.: Center for Applied Linguistics.  
Windfuhr, Gernot L. 1997. Persian phonology. In Kaye, Alan S. and Daniels, Peter T. (eds.), Phonologies of Asia and Africa, 675–690. Winona Lake, Indiana: Eisenbrauns.  

Majidi, Mohammad-Reza and Ternes, Elmar. 1991. Illustrations of the IPA: Persian (Farsi). Journal of the International Phonetic Association 21. 96–98. Cambridge University Press.  
  - Page 97 : Trill r  
  - Page 98 : /r/ varies between [ɾ] and [ɹ]  

Wikipedia :     
https://en.wikipedia.org/wiki/Iranian_Persian  
https://en.wikipedia.org/wiki/Persian_phonology  
  - Trill r 					  
  - Tap ɾ  
  - The flap /ɾ/ has a trilled allophone [r] at the beginning of a word;[16] otherwise, they contrast between vowels wherein a trill occurs as a result of gemination (doubling) of [ɾ], especially in loanwords of Arabic origin. Only [ɾ] occurs before and after consonants; in word-final position, it is usually a free variation between a flap or a trill when followed by a consonant or a pause, but flap is more common, only flap before vowel-initial words. An approximant [ɹ] also occurs as an allophone of /ɾ/ before /t, d, s, z, ʃ, l, ʒ/; [ɹ] is sometimes in free variation with [ɾ] in these and other positions, such that فارسی ('Persian') is pronounced [fɒːɹˈsiː] or [fɒːɾˈsiː] and سقرلات ('scarlet') [sæɣeɹˈlɒːt] or [sæɣeɾˈlɒːt]. /r/ is sometimes realized as a long approximant [ɹː].   

Mahootian, Shahrzad (1997). Persian. London: Routledge. pp. 287, 292, 303, 305. ISBN 0-415-02311-4.  
  - Page 287 : Liquid r
  - Page 292 : Voiced nonsyllabic /r/ is made by flapping the tip of the tongue forward against the alveolar ridge. There is allophonic variation, with more trill at the beginning of a word than in the middle or at the end. Both /l/ and /r/ occur word-initially and word-finally, syllable-initially and syllable-finally as well as morpheme-initially
and morpheme-finally. Liquids also occur in word-medial, syllablemedial and morpheme-medial positions 

Glottologue :    
https://glottolog.org/resource/languoid/id/west2369  

Autres sources : NA

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="farsi",1,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="farsi",1,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="farsi",1,trill_occ))
```

## finnish

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="finnish","finn1318",glottocode))
```

Phoible :   
https://phoible.org/languages/finn1318  
3 inventaires  
  - SPA r allophones r; et rː allophones rː   ; contrast r/l
  - UPSID	r̪|r pas d'allophones  ; contrast r̪|r/l̪
  -EA r̺ pas d'allophones    ; contrast r̺/l̺

Suomi, Kari and Toivanen, Juhani and Ylitalo, Riikka. 2010. Finnish sound structure: Phonetics, phonology, phonotactics and prosody. Puhe ja kieli 30. 121–124.  
  - Page 30:  
  The rhotic /r/  
  It is not straightforward to determine the main allophone of /r/. Traditionally, it  
  has been stated that the main allophone is a trill, but recent investigations suggest  
  that the tap pronunciation ([ɾ]) is much more common than has been previously    
  presumed. Thus Mustanoja & O’Dell (2007) observed, in two corpora (colloquial  
  speech in Tampere and news broadcasts of the national broadcasting company)  
  that a great majority of the single /r/ productions in word-medial intervocalic  
  position were taps (75% and 90% in the two corpora, respectively). It seems   
  however that in other positions, e.g. word-initially, a trill realisation is more  
  common than the tap. Also the quantity opposition has an effect on the realisation  
  of /r/. Thus, as just mentioned, a single /r/ (as in paras ‘best’) is often realised as  
  [ɾ], but double /rr/ (as in parras ‘edge’) always as [r:], often with several closure  
  periods. On the whole, it seems legitimate to conclude that the main allophone of  
  /r/ is [r].  
  
  In addition, /r/ has an alveolar fricative allophone, [®3]. As was mentioned  
  above in discussing the allophones of /s/, the sequence /sr/ does not occur inside  
  native uncompounded words. A very likely reason for this avoidance is that the  
  sequence of the main allophones of the two phonemes, i.e. [sr], a sequence  
  consisting of a sibilant and a trill, is difficult to pronounce even for a native  
  phonetician. However, the phoneme sequence /sr/ is common across a word  
  boundary, and across a boundary between the parts of compound words. One way  
  of avoiding the pronunciation difficulty in e.g. Israel is that the allophonic  
  sequence [s®3] is chosen. Another way, mentioned earlier, is to choose the  
  allophonic sequence [xr]. Thus, the difficult sequence sibilant + trill is avoided  
  either by replacing the trill by a fricative rhotic, or by replacing the sibilant by a  
  non-sibilant fricative. The sequence [x®3] is not attested. The following account of  
  the allophones of /r/ is somewhat tentative:  
  [rule]  
  - Page 31 : In addition to these allophones proper, /r/ has more marginal realisations that  
  occur in some speakers’ idiolects, realisations that are not dialectal properties nor  
  due to segmental context. These include [®] and [“]. There is very little areal  
  variation in the regular allophones of /r/, but in the Tampere area their place of  
  articulation is further back than in other dialects, probably apical retroflex  
  (Kuronen 2000).  
  - Page 57 : /l/ and /s/ cannot be followed by /r/. Presumably both restrictions have a  
  similar articulatory motivation: /r/ requires an activity of the tongue blade  
  that is difficult to accomplish after /l/ and /s/. Exceptions: loanwords like  
  Kilroy, Israel. The reverse sequences /rl/ and /rs/ in turn are common. Recall  
  from above that the sequence /sr/ (which is common across a word boundary  
  and across the boundary between the components of compound words) has  
  the alternative pronunciations [xr] and [s®3 ]. Thus in the /sr/ sequence both  
  phonemes are realised, but in such a way that a sibilant is not followed by a   
  trill. In fact, this is a kind of phonetic sandhi restriction: the phonemic   
  affiliations of the consonants do not change, but if the occurring allophone of  
  one of the consonants is a sibilant or a trill, then the other consonant must be  
  realised by an allophone that does not occur elsewhere.  


Je n'ai pas regardé les sources plus vieilles de PHOIBLE :  
Austerlitz, Robert. 1967. The Distributional Identification of Finnish Morphophonemes. Language 43. 20–33.  
Lehtinen, M. 1964. Basic Course in Finnish. Bloomington: Indiana University.  
Hammarberg, R. 1974. Another Look at Finnish Consonant Gradation. Soviet Finno-Ugric Studies 10. 171–178.  
Harms, Robert T. 1966. Review Of: Proto-Finnic Final Consonants, by T. Itkonen. Language 42. 825–831.  
Harms, Robert T. 1964. Finnish Structural Sketch. Bloomington: Indiana University Press.  

Wikipedia :  
https://en.wikipedia.org/wiki/Finnish_language  
https://en.wikipedia.org/wiki/Finnish_phonology  
  - Trill	r  

Glottologue :  
https://glottolog.org/resource/languoid/id/finn1318  

Autres sources : NA  

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="finnish",1,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="finnish",1,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="finnish",1,trill_occ))
```

## french

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="french","stan1290",glottocode))
```

Phoible :     
https://phoible.org/languages/stan1290  
4 inventaires :  
  - SPA r allophones r̥ r; ʀ allophones ʀ ʁ̞ ʀ̥  
  - UPSID  ʀ pas d'allophone  
  - UZ ʁ allophones	ʁ ʁ̞ ʁ̥ ʀ r  
  - EA ʁ pas d'allophone  
  
Sten, H. 1963. Manuel de Phonetique Francaise. Copenhagen: Munksgaard.  

Fougeron, Cécile and Smith, Caroline L. 1993. Illustrations of the IPA: French. Journal of the International Phonetic Association 23. 73–76. Cambridge University Press.
  - Page 74 : Fricative ʁ  
              French has one rhotic, whose production varies considerably among speakers and phonetic contexts. The speaker presented here uses a uvular fricative [ʁ] that is
sometimes reduced to an approximant [ʁ̞], particularly in final position; it may also be devoiced (for examples see the transcribed text), and can be reduced to zero in some
word-final positions. For other speakers, a uvular trill [ʀ] is also fairly common, and an apical trill [r] occurs in some dialects. Vowels are often lengthened before this segment.  

Wikipedia :   
https://en.wikipedia.org/wiki/French_language    
  - Fricative ʁ    
  French has one rhotic whose pronunciation varies considerably among speakers and phonetic contexts. In general, it is described as a voiced uvular fricative, as in [ʁu] roue, "wheel". Vowels are often lengthened before this segment. It can be reduced to an approximant, particularly in final position (e.g., fort), or reduced to zero in some word-final positions. For other speakers, a uvular trill is also common, and an apical trill [r] occurs in some dialects.     
  
Glottologue :    
https://glottolog.org/resource/languoid/id/stan1290    

Autres sources : NA  

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="french",1,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="french",0,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="french",1,trill_occ))
```

## georgian

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="georgian","nucl1302",glottocode))
```

Phoible :   
https://phoible.org/languages/nucl1302  
4 inventaires  
  SPA r	allophones r r̥ r ɾ (ʁ allophones ʁ χ͉ ʁ)   ; contrast r/lˠ
  UPSID	r̪|r pas d'allophones   ; contrast r̪|r/l̪ˠ|lˠ
  UZ ɾ allophones ɾ r ɾ̩ r̩ l (l	allophones	l l̩ ɾ)  ; contrast r/l
  EA ɾ pas d'allophones (ʁ pas d'allophones)   ; contrast  ɾ/ɫ

Shosted, Ryan K. and Chikovani, Vakhtang. 2006. Illustrations of the IPA: Standard Georgian. Journal of the International Phonetic Association 36. 255–264. Cambridge University Press.  
  - Page 255 : Tap/Trill ɾ/r
  - Page 261 : Although it has been claimed that the Georgian /r/ may be ‘rolled’, this does not seem to be the case for the present speaker (Hewitt 2005: 5). In our tokens, the average closure duration for an intervocalic /r/ is quite short, on the order of 20 ms, and there is little evidence for sustained trilling in the audio signal. While the two brief reductions in oral flow observed in figure 5 may be indicative of a two-strike trill, it is probably safe to say that taps, not multiple-strikes, are the pronunciation norm in Standard Georgian.


Aronson, Howard I. 1990. Georgian: A Reading Grammar. Corr. ed. edn. Columbus, Ohio: Slavica Publishers, Indiana University. 526pp. (Corrected version (original from 1982).)  
  - r. Georgian r is generally formed by a single flap of the tip of the tongue to the alveolar ridge [ɾ]. It is very similar to the r in Spanish pe/r/o 'but' or the Russian r in [Word in Russian]   'break'. Between voiceless consonants it may become voiceless or drop completely.  
Vogt, Hans. 1971. Grammaire de la langue géorgienne. Oslo: The Institute for Comparative Research in Human Culture.  
  - sonantes r  
  - Page 12-13 : La sonante r est un r apico-alvéolaire roulé. Le nombre des battements n'excède guère deux ou trois après consonne devant voyelle, et peut se réduire à un seul en position finale. Elle peut s'assourdir en position finale après consonne [...] et entre deux consonnes sourdes [...].  

Wikipedia :    
https://en.wikipedia.org/wiki/Georgian_language#Consonants  
  - Vibrant r  
  /r/ is realised as an alveolar tap [ɾ]  

Avec comme source : Shosted, Ryan K.; Vakhtang, Chikovani (2006), "Standard Georgian"  
http://accent.gmu.edu/browse_native.php?function=detail&languageid=23  
  - tap or flap ɾ  
  A partir de Phonologies of Asia and Africa Edited by Alan S. Kaye and Peter T. Daniels  

Glottologue :  
https://glottolog.org/resource/languoid/id/nucl1302  

Autres sources : NA  

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="georgian",1,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="georgian",0,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="georgian",1,trill_occ))
```

## german

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="german","stan1295",glottocode))
```

Phoible :   
https://phoible.org/languages/stan1295  
4 inventaires  
  - SPA : ʀ	allophones ʌ ʌ̯ ʀ r   ; contrast ʀ/l
  - UPSID : ʀ pas d'allophone   ; contrast ʀ/l
  - UZ : ʁ	pas d'allophone	ʁ ʔ ʁ̥ χ ɐ  ; contrast ʁ/l
  - EA : ʀ pas d'allophone  ; contrast ʀ/l
  
Moulton, William G. 1962. The Sounds of English and German. Chicago: University of Chicago Press.  
Werner, Otmar. 1972. Phonemik des Deutschen. (Sammlung Metzler, 108.) Stuttgart: J.B. Metzler Verlagsbuchhandlung.  
Philipp, Marthe. 1974. Phonologie des Deutschen. Stuttgart: W. Kohlhammer.  

Kohler, Klaus. 1990. Illustrations of the IPA: German. Journal of the International Phonetic Association 20. 48–50. Cambridge University Press.  
  - Page 48 : Fricative ʁ
              ʁ Rasse 'race'
  - Page 49 : /ʁ/ can be an approximant intervocalically (e.g. Herren ˈgentlemenˈ); after voiceless plosives and fricatives, especially those within the same word, it is devoiced (in e.g. trat ˈkickedˈ it is completely voiceless [χ]]); postvocalically before a consonant or word-final it is vocalized to [ɐ], which results in diphthongs (e.g. [ˈhaɐt] hart ˈhardˈ, [ˈoɐ] Ohr ˈearˈ, see
the vowel chart below); the ending -er is realised as [ɐ] (e.g. [ˈbʊtɐ] Butter ˈbutterˈ); the place of articulation of the consonant varies from uvular in e.g. rot ˈredˈ to velar in e.g. treten ˈkickˈ, depending on back or front vowel contexts./ can be an approximant intervocalically (e.g. Herren ˈgentlemenˈ); after voiceless plosives and fricatives, especially those within the same word, it is devoiced (in e.g. trat ˈkickedˈ it is completely voiceless ft]); postvocalically before a consonant or word-final it
is vocalized to [B], which results in diphthongs (e.g. [ˈham] hart ˈhardˈ, [ˈoe] Ohr ˈearˈ, see the vowel chart below); the ending -er is realised as [B] (e.g. [ˈbutB] Butter ˈbutterˈ); the place of articulation of the consonant varies from uvular in e.g. rot ˈredˈ to velar in e.g. treten ˈkickˈ, depending on back or front vowel contexts.

Wikipedia :   
https://en.wikipedia.org/wiki/German_language
  - Liquid r 	

https://en.wikipedia.org/wiki/Standard_German_phonology
  - Liquid r 	
  - /r/ can be uvular, alveolar or even dental, a consonant or a semivowel, see below.
  - /r/ has a number of possible realizations:
    Voiced apical coronal trill [r̺],[58][59][60] either alveolar (articulated with the tip of the tongue against the alveolar ridge),[58][59][60] or dental (articulated with the tip of the tongue against the back of the upper front teeth).[58]
    Distribution: Common in the south (Bavaria and many parts of Switzerland and Austria), but it is also found in some speakers in central and northern Germany, especially the elderly. It is also one of possible realizations of /r/ in the Standard Austrian accent, but a more common alveolar realization is an approximant [ɹ]. Even more common are uvular realizations, fricatives [ʁ ~ χ] and a trill [ʀ].[61]
    Voiced uvular trill [ʀ],[58][59][62][63] which can be realized as voiceless [ʀ̥] after voiceless consonants (as in treten).[59] According to Lodge (2009) it is often a flap [ʀ̆] intervocalically (as in Ehre).[64]
    Distribution: Occurs in some conservative varieties - most speakers with a uvular /r/ realize it as a fricative or an approximant.[65] It is also one of possible realizations of /r/ in the Standard Austrian accent, but it is less common than a fricative [ʁ ~ χ].[61]
    Dorsal continuant, about the quality of which there is not a complete agreement:
    Krech et al. (2009) describe two fricative variants, namely post-palatal [ɣ˖] and velar [ɣ]. The post-palatal variant appears before and after front vowels, while the velar variant is used in all other positions.[66]
    Morciniec & Prędota (2005) describe it as voiced post-velar fricative [ʁ̟].[67]
    Mangold (2005) and Kohler (1999) describe it as voiced uvular fricative [ʁ];[58][68]
    Mangold (2005) states that "with educated professional radio and TV announcers, as with professional actors on the stage and in film, the [voiced uvular] fricative [realization of] /r/ clearly predominates."[58]
    In the Standard Austrian accent, the uvular fricative is also the most common realization, although its voicing is variable (that is, it can be either voiced [ʁ] or voiceless [χ]).[61]
    Kohler (1999) writes that "the place of articulation of the consonant varies from uvular in e.g. rot ('red') to velar in e.g. treten ('kick'), depending on back or front vowel contexts." He also notes that [ʁ] is devoiced after voiceless plosives and fricatives, especially those within the same word, giving the word treten as an example. According to this author, [ʁ] can be reduced to an approximant in an intervocalic position.[69]
    Ladefoged & Maddieson (1996) describe it as a uvular fricative [ʁ] or approximant [ʁ̞]. The latter is less likely to occur word-initially.[70]
    Distribution: Almost all areas apart from Bavaria and parts of Switzerland.
    Near-open central unrounded vowel [ɐ] is a post-vocalic allophone of (mostly dorsal) varieties of /r/. The non-syllabic variant of it is not always near-open or central; it is similar to either [ɑ] or [ə], depending on the environment.[67]
    distribution: Widespread, but less common in Switzerland.

Glottologue :  
https://glottolog.org/resource/languoid/id/stan1295  

Autres sources : NA  
```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="german",1,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="german",0,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="german",1,trill_occ))
```

## greek

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="greek","mode1248",glottocode))
```

Phoible :     
https://phoible.org/languages/mode1248  
5 inventaires:  
  - SPA ɾ allophone ɾ   ;  contrast ɾ/l
  - UPSID ɾ pas d'allophone    ; contrast ɾ/l̪
  - UZ 2186	rː allophone rː; ɾ allophones ɾ rː    ; contrast  ɾ/l
  - UZ 2187 ɾ allophones ɾ ɹ    ; contrast ɾ/l
  - EA ɾ̠ pas d'allophone  ; contrastɾ ɾ̠/l
 
Kaisse, Ellen. 1976. Stress Melodies and a Fast Speech Rule in Modern Greek. NELS VI. 165–175.  
Newton, Brian. 1972. The Generative Interpretation of Dialect: A Study of Modern Greek Phonology. (Cambridge Studies in Linguistics, 8.) Cambridge: Cambridge University Press.  
Kaisse, Ellen. 1975. A Mor- or Less-pheme: A Case of Delexicalization in Modern Greek. Paper delivered at the 1975 Winter Meeting of the LSA, San Francisco.  
Householder, Fred W. and Kazasis, Kostas and Koutsoudas, Andreas. 1964. A Reference Grammar of Literary Dhimotiki. Bloomington, Indiana: Indiana University Press.  
Pring, J. T. 1967. A Grammar of Modern Greek on a Phonetic Basis. London: Hodder & Stoughton.  
Arvaniti, Amalia. 2007. Greek phonetics: The state of the art. Journal of Greek Linguistics 8. 97–208. Citeseer.  

Arvaniti, Amalia. 1999. Illustrations of the IPA: Standard Modern Greek. Journal of the International Phonetic Association 29. 167–172/1–6. Cambridge University Press.  
  - Page 167 : Tap ɾ  
  - Page 168-169 : Rhotics. Greek has one rhotic sound. This is pronounced as an alveolar tap [ɾ] word-initially and intervocalically, or when it is followed by another consonant; in fast speech it may be pronounced as an alveolar approximate In clusters In which it is preceded by a plosive or a fricative it is pronounced as a short trill.  
  
Wikipedia :     
https://en.wikipedia.org/wiki/Modern_Greek  
https://en.wikipedia.org/wiki/Modern_Greek_phonology  
  Tap /ɾ/ ρ  
 
  The only Greek rhotic /r/ is prototypically an alveolar tap [ɾ], often retracted ([ɾ̠]). It may be an alveolar approximant [ɹ] intervocalically, and is usually a trill [r] in clusters, with two or three short cycles.[8]  
    [8] Arvaniti, Amalia (2007). "Greek Phonetics: The State of the Art" (PDF). Journal of Greek Linguistics. 8: 97–208. CiteSeerX 10.1.1.692.1365. doi:10.1075/jgl.8.08arv. Archived from the original (PDF) on 2013-12-11.  


Glottologue :    
https://glottolog.org/resource/languoid/id/mode1248  

Autres sources : NA  

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="greek",1,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="greek",0,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="greek",1,trill_occ))
```

## hebrew

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="hebrew","hebr1245",glottocode))
```

Phoible :   
https://phoible.org/languages/hebr1245  
3 inventaires  
  - SPA	ʁ allophones ʀ ʁ   ; contrast ʁ/l
  - UZ 	r	allophones	r ɾ; 	χ	allophones	χ    ; contrast r/l
  - EA  ʁ pas d'allophones    ; contrast ʁ/l
 
Cohen, David and Zafrani, Haim. 1968. Grammaire de l'Hebreu Vivant. Paris: Presses Universitaives de France. -> Pas cherché  

Chayen, M. J. 1973. The Phonetics of Modern Hebrew. (Janua Linguarum, Series Practica, 163.) The Hague: Mouton.  
  - Page 24 : The voiced member of this pair, /r/, is sometimes trilled with the back of the tongue tapping rapidly against the soft palate.  
              Most recordings however show a consistent fricative /r/ with the trill rare.  
              Arabic speakers have a tongue-trilled apical /r/ which is recommended by official bodies (the radio).[10]  
              
              [9] I have departed from IPA notation here and use [r] to represent fricative velar and [R] for trilled velar.  
              [10] Y.G. Gumpertz (1953) quotes Sa'adia Gaon (10th centenry) and other early grammarians to show that Hebrew had both alveolar and uvular /r/, the former following dental consonants.  
  - Page 25 : Both /x/ and /r/ have unrestricted distribution   
  
Laufer, Asher. 1990. Illustrations of the IPA: Hebrew. Journal of the International Phonetic Association 20. 40–43. Cambridge University Press.  
  - Page 41 : Trill r  
  - Page 42 : Usually the Oriental-Hebrew speakers pronounce /r/ as an alveolar trill (or flap) [ɾ]; in the Non-Oriental dialect it is usually a ucular approximant [ʁ] (still, some Non-Oriental speakers pronounce it as a trill).  
  
Bolozky, Shmuel. 1997. Israeli Hebrew Phonology. In Kaye, Alan S. and Daniels, Peter T. (eds.), Phonologies of Asia and Africa, 287–312. Winona Lake, Indiana: Eisenbrauns. -> Pas cherché  

Wikipedia :   
https://en.wikipedia.org/wiki/Modern_Hebrew  
  - Obstruents Fricative ɣ~ʁ  
  - Commonly transcribed /r/. This is usually pronounced as a uvular fricative or approximant [ʁ] or velar fricative [ɣ], and sometimes as a uvular [ʀ] or alveolar trill [r] or alveolar flap [ɾ], depending on the background of the speaker.  

https://en.wikipedia.org/wiki/Hebrew_language  
  - Most Israelis today also merge /ʕ ħ/ with /ʔ χ/, do not have contrastive gemination, and pronounce /r/ as a uvular fricative [ʁ] or a voiced velar fricative [ɣ] rather than an alveolar trill, because of Ashkenazi Hebrew influences.  

Glottologue :  
https://glottolog.org/resource/languoid/id/hebr1245  

Autres sources : NA  

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="hebrew",1,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="hebrew",0,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="hebrew",1,trill_occ))
```

## hindi

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="hindi","hind1269",glottocode))
```

Phoible :   
https://phoible.org/languages/hind1269
5 inventaires :
  - SPA r allophones r ; rː allophones rː ; ɽ̃ allophones 	ɽ̃ ɽ̃ ɽ̃ ; ʁ allophone ʁ ; ɖ allophones ɖ ɖ ɳɖ ɽ ; ɖ̤ allophones ɖ̤ ɳɖ̤ ɽ̤ ; contrast r/l
  - UPSID 	r̪|r pas d'allophones ; contrast 	r̪|r/l̪|l
  - RA ɽ pas d'allophone ; 	ɽ̤ pas d'allophone  ; ɾ pas d'allophone  ; contrast ɾ/l
  - UZ 	ɾ allophones ɾ r ; rː allophone rː ;  ɽ allophone ɽ ; ɽ̤ allophone ɽ̤  ; contrast	rː/l
  - EA r pas d'allophone ; ɽ pas d'allophone ; ɽ̤ʱ pas d'allophone  ; contrast r/l

Pinnow, Heinz-Jürgen. 1972. "Uber die Vokale im Hindi. Zeitschrift fur Phonetik und Allgemeine Sprachwissenschaft 7. 43–53.
Kelkar, Ashok Ramchandra. 1968. Studies in Hindi-Urdu. Volume 1: Introduction and Word Phonology. Poona: Postgraduate and Research Institute, Deccan College.
Vermeer, Hans J. and Sharma, Aryendra. 1966. Hindi-Lautlehre mit Einführung in die Devnagari-Schrift. Heidelberg: Julius Groos Verlag.
Harms, Robert T. 1969. Review of Kelkar, A.R. Studies in Hindi-Urdu I. Language 45. 913–927.
Kaye, Alan S. 1997. Hindi-Urdu phonology. In Kaye, Alan S. and Daniels, Peter T. (eds.), Phonologies of Asia and Africa, 637–652. Winona Lake, Indiana: Eisenbrauns.
Ohala, Manjari. 1983. Aspects of Hindi Phonology. Delhi: Motilal Banarsidass. (Via https://archive.org/details/aspectsofhindiph0000ohal)
  - Page 2 : Tap  Dental and Alveolar r r:
             Flaps  Retroflex voiced         r.  (r̃.)
                              breathy-voiced r.^h
  - Page 4 : [Ça parle de r. qui ressemble à un [n]]
  - Page 5 : [Mention d'un r syllabique -> aucune idée de la phonétique]
  - Page 6 : [Ça parle de r. avec d.]
  
Wikipedia :   
https://en.wikipedia.org/wiki/Hindi
https://en.wikipedia.org/wiki/Hindustani_phonology
  - Tap/Trill 	unaspirated 		r 	ɽ 				
                aspirated 			ɽʱ
  - /ɽ/ is lateral [ɺ̢] for some speakers.
  - /r/ is usually flapped or trilled.[24] In intervocalic position, it may have a single contact and be described as a flap [ɾ],[25] but it may also be a clear trill, especially in word-initial and syllable-final positions, and geminate /rː/ is always a trill in Arabic and Persian loanwords, e.g. zarā [zəɾaː] (ज़रा – ذرا 'little') versus well-trilled zarrā [zəraː] (ज़र्रा – ذرّہ 'particle').[4]
  [24] Nazir Hassan (1980) Urdu phonetic reader, Omkar Nath Koul (1994) Hindi Phonetic Reader, Indian Institute of Language Studies; Foreign Service Institute (1957) Hindi: Basic Course
  [25] "r is a tip dental trill, and often has but one flap", Thomas Cummings (1915) An Urdu Manual of the Phonetic, Inductive Or Direct Method
  [4] Ohala, Manjari (1999), "Hindi", in International Phonetic Association (ed.), Handbook of the International Phonetic Association: a Guide to the Use of the International Phonetic Alphabet, Cambridge University Press, pp. 100–103, ISBN 978-0-521-63751-0

Glottologue :  
https://glottolog.org/resource/languoid/id/hind1269

Autres sources :  
Ohala, M. (1994). Hindi. Journal of the International Phonetic Association, 24(1), 35‑38. https://doi.org/10.1017/S0025100300004990
  - Page 35 : Tap or Flap ɾ   ɽ
                              ɽ^h
  - Page 36 : All consonants can also occur with distinctive length (i.e. as 'geminates') except the following: [b^h, ɽ, ɽ^h, h].
  - Page 37 : The geminate [r] is always an alveolar trill [r]
  
```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="hindi",1,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="hindi",0,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="hindi",1,trill_occ))
```

## hungarian

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="hungarian","hung1274",glottocode))
```

Phoible :   
https://phoible.org/languages/hung1274
4 inventaires :
  - SPA	r allophone r ;	rː allophone rː ; contrast r/l 
  - UPSID	r̪ pas d'allophone ; contrast r̪/l̪
  - UZ 	r̪ allophone 	r̪ ; 	r̪ː allophone	r̪ː ; contrast r̪/l̪
  - EA ɾ pas d'allophone ; contrast ɾ/l

Hall, Robert Anderson. 1938. An Analytical Grammar of the Hungarian Language. Language 14. 9–113.
Hall, Robert Anderson. 1944. Hungarian Grammar. (Language Monographs, 21.) Baltimore: Linguistic Society of America.
Kalman, Bela. 1972. Hungarian Historical Phonology. In Benkö, Lorand and Imre, Samu (eds.), The Hungarian Language. The Hague: Mouton.
Banhidi, Zoltan and Jokay, Zoltan and Szabo, Denes. 1965. Lehrbuch der ungarischen Sprache. Budapest: Publishing House for Textbooks.
Gósy, Mária. 2004. Fonetika, a beszéd tudománya. Osiris.

Szende, Tamás. 1994. Illustrations of the IPA: Hungarian. Journal of the International Phonetic Association 24. 91–94. Cambridge University Press.
  - Page 91 : Trill r
              r ro: ró 'carve'
  - Page 93 : /r/ is a trill

Wikipedia :   
https://en.wikipedia.org/wiki/Hungarian_language
  - Trill 			r 	
  - A single 'r' is pronounced as an alveolar tap (akkora 'of that size'), but a double 'r' is pronounced as an alveolar trill (akkorra 'by that time'), like in Spanish and Italian. 

https://en.wikipedia.org/wiki/Hungarian_phonology
  - Trill 			r
  - /r/ 	ró 	/roː/ 'carve' 
  - /l/ assimilates to a following /r/ (e.g. balra [ˈbɒrːɒ], 'to the left').[27]
  
  [27] Vago, Robert M. (1980), The Sound Pattern of Hungarian, Washington, D.C.: Georgetown University Press

Glottologue :  
https://glottolog.org/resource/languoid/id/hung1274

Autres sources : NA  

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="hungarian",1,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="hungarian",1,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="hungarian",1,trill_occ))
```

## icelandic

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="icelandic","GLOTTOCODE",glottocode))
```

Phoible :   

Wikipedia :   

Glottologue :  

Autres sources :  

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="icelandic",10,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="icelandic",10,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="icelandic",10,trill_occ))
```

## inari

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="LANGUAGE","GLOTTOCODE",glottocode))
```

Phoible :   

Wikipedia :   

Glottologue :  

Autres sources :  

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="LANGUE",10,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="LANGUE",10,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="LANGUE",10,trill_occ))
```

## indonesian

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="LANGUAGE","GLOTTOCODE",glottocode))
```

Phoible :   

Wikipedia :   

Glottologue :  

Autres sources :  

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="LANGUE",10,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="LANGUE",10,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="LANGUE",10,trill_occ))
```

## irish

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="LANGUAGE","GLOTTOCODE",glottocode))
```

Phoible :   

Wikipedia :   

Glottologue :  

Autres sources :  

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="LANGUE",10,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="LANGUE",10,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="LANGUE",10,trill_occ))
```

## italian

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="LANGUAGE","GLOTTOCODE",glottocode))
```

Phoible :   

Wikipedia :   

Glottologue :  

Autres sources :  

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="LANGUE",10,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="LANGUE",10,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="LANGUE",10,trill_occ))
```

## japanese

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="LANGUAGE","GLOTTOCODE",glottocode))
```

Phoible :   

Wikipedia :   

Glottologue :  

Autres sources :  

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="LANGUE",10,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="LANGUE",10,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="LANGUE",10,trill_occ))
```

## konkani

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="LANGUAGE","GLOTTOCODE",glottocode))
```

Phoible :   

Wikipedia :   

Glottologue :  

Autres sources :  

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="LANGUE",10,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="LANGUE",10,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="LANGUE",10,trill_occ))
```

## korean

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="LANGUAGE","GLOTTOCODE",glottocode))
```

Phoible :   

Wikipedia :   

Glottologue :  

Autres sources :  

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="LANGUE",10,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="LANGUE",10,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="LANGUE",10,trill_occ))
```

## kurdish

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="LANGUAGE","GLOTTOCODE",glottocode))
```

Phoible :   

Wikipedia :   

Glottologue :  

Autres sources :  

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="LANGUE",10,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="LANGUE",10,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="LANGUE",10,trill_occ))
```

## latin

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="LANGUAGE","GLOTTOCODE",glottocode))
```

Phoible :   

Wikipedia :   

Glottologue :  

Autres sources :  

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="LANGUE",10,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="LANGUE",10,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="LANGUE",10,trill_occ))
```

## latvian

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="LANGUAGE","GLOTTOCODE",glottocode))
```

Phoible :   

Wikipedia :   

Glottologue :  

Autres sources :  

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="LANGUE",10,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="LANGUE",10,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="LANGUE",10,trill_occ))
```

## limburgish

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="LANGUAGE","GLOTTOCODE",glottocode))
```

Phoible :   

Wikipedia :   

Glottologue :  

Autres sources :  

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="LANGUE",10,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="LANGUE",10,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="LANGUE",10,trill_occ))
```

## macedonian

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="LANGUAGE","GLOTTOCODE",glottocode))
```

Phoible :   

Wikipedia :   

Glottologue :  

Autres sources :  

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="LANGUE",10,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="LANGUE",10,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="LANGUE",10,trill_occ))
```

## malay

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="LANGUAGE","GLOTTOCODE",glottocode))
```

Phoible :   

Wikipedia :   

Glottologue :  

Autres sources :  

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="LANGUE",10,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="LANGUE",10,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="LANGUE",10,trill_occ))
```

## mingrelian

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(glottocode = ifelse(Languages=="LANGUAGE","GLOTTOCODE",glottocode))
```

Phoible :   

Wikipedia :   

Glottologue :  

Autres sources :  

```{r}
languages_data <- languages_data %>%
    dplyr::mutate(r_l_distinction = ifelse(Languages=="LANGUE",10,r_l_distinction))

languages_data <- languages_data %>%
    dplyr::mutate(trill_real = ifelse(Languages=="LANGUE",10,trill_real))

languages_data <- languages_data %>%
    dplyr::mutate(trill_occ = ifelse(Languages=="LANGUE",10,trill_occ))
```

## nepali
## norwegian

"norwegian"      "pashto"         "polish"         "portuguese"     "romanian"       "russian"        "saami"          "sanskrit"      
"serbian"        "silesian"       "slovak"         "sotho"          "spanish"        "swahili"        "swazi"          "swedish"       
"swissgerman"    "tamil"          "thai"           "turkish"        "ukrainian"      "vietnamese"     "wolof"          "xhosa"         
"zhuang"         "zulu"

##


##


##
