---
title: "R/L across languages: Preparation"
author: "Aleksandra Ćwiek"
date: "2023-10-26"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 4
    toc_float: yes
    df_print: paged
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: '3'
  html_notebook:
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data preparation

```{r pckgs, message=FALSE, warning=FALSE}
library(tidyverse)

# Set the script's path as working directory
parentfolder = rstudioapi::getActiveDocumentContext()$path 
setwd(dirname(parentfolder))
parentfolder <- getwd()

models        <- paste0(parentfolder, '/models/')
plots         <- paste0(parentfolder, '/plots/')
data          <- paste0(parentfolder, '/data/')
```

Load the data

```{r load data, message=FALSE, warning=FALSE}
field <- read.csv(paste0(data, "fieldwork-responses-rest.csv"), sep = ";")

web <- read.csv(paste0(data, "r_l_web.csv"), sep = ",")

languages_data <- read.csv(paste0(data, "languages_data.csv"), sep = ",")
```

# Online experiment

## Data cleaning & exclusions

Languages are stored in the format "AL", "EN" etc. Let's process the "experiment" column so that it has only those strings:

```{r extract languages, message=FALSE, warning=FALSE}
web <- mutate(web,
              Language = str_extract(experiment, '[A-Z]{2,2}'))
```

Just in case some session IDs recur across the different experiments, let's create a unique identifier column by pasting languages and session IDs together:

```{r create identifier, message=FALSE, warning=FALSE}
web <- mutate(web,
              ID = str_c(Language, '_', session))
```

Exclude those that have 'l1' == 'q' or 'b' which is Ola's code for her own test runs.

```{r filtering tests, message=FALSE, warning=FALSE}
web <- filter(web,
              l1 != 'q')
web <- filter(web,
              l1 != 'b')
web <- filter(web,
              firstlanguage != 'q')
```

How many participants is that before exclusions (because of playbacks etc.)?

```{r how many, message=FALSE, warning=FALSE}
length(unique(web$ID))
```

Check number of playbacks:

```{r playcheck, message=FALSE, warning=FALSE}
table(web$playbacks)
```

Exclude 0 playbacks (didn't listen to the sound):

```{r playrid, message=FALSE, warning=FALSE}
web <- filter(web, playbacks != 0)
```

Check ages:

```{r agecheck, message=FALSE, warning=FALSE}
table(web$participantage)
```

Get rid of everybody below 18:

```{r agerid, message=FALSE, warning=FALSE}
web <- filter(web, participantage > 17)
```

Check input values:

```{r check input, message=FALSE, warning=FALSE}
table(web$inputvalue)
```

Get rid of the responses that are neither "r" nor "l":

```{r exclude false input, message=FALSE, warning=FALSE}
web <- filter(web, !inputvalue %in% c('klein', 'ogień'))
```

## Process L1 information:

Create a table with one data point per participant with L2 information:

```{r create l2 info, message=FALSE, warning=FALSE}
L2_info <- filter(web, !duplicated(ID)) %>% select(ID, Language, l1, l2)
```

Get rid of trailing spaces:

```{r remove trailing, message=FALSE, warning=FALSE}
L2_info <- mutate(L2_info,
                  l1 = str_trim(l1))
```

That one Chinese person who wrote "hi" in the language field probably knows English given that they used this word... this is the more conservative choice (working against our hypothesis):

```{r adapt l2 info, message=FALSE, warning=FALSE}
L2_info[L2_info$l1 == 'hi', ]$l1 <- 'chinese'
L2_info[which(L2_info$l2 == 'hi'), ]$l2 <- 'english'
```

<!-- Perform replacements — for this, we take whatever is the person's first-listed language as their primary L1 (this only applies to a few cases anyway). In a few cases where there is overlap with another language of which we have multiple speakers in the sample, the assignment will be done this way, e.g., "ukrainian, russian" -> "russian"

```{r replacements l2 info_old, message=FALSE, warning=FALSE, eval=FALSE}
L2_info <- mutate(L2_info,
                  l1 = str_replace(l1, 'isizulu', 'zulu'),
                  l1 = str_replace(l1, 'turkce', 'turkish'),
                  l1 = str_replace(l1, 'swedish, korean', 'swedish'),
                  l1 = str_replace(l1, 'french, spanish', 'french'),
                  l1 = str_replace(l1, 'ukrainian, russian', 'russian'),
                  l1 = str_replace(l1, 'portuguese, french, danish', 'portuguese'),
                  l1 = str_replace(l1, 'finnish, estonian', 'finnish'),
                  l1 = str_replace(l1, 'albanianja', 'albanian'),
                  l1 = str_replace(l1, 'armenian, russian', 'armenian'),
                  l1 = str_replace(l1, 'swissgerman', 'german'),
                  l1 = str_replace(l1, 'austriangerman', 'german'),
                  l1 = str_replace(l1, 'german, turkish', 'german'),
                  l1 = str_replace(l1, 'german, english', 'german'),
                  l1 = str_replace(l1, 'konkani, english', 'english'),
                  l1 = str_replace(l1, 'español', 'spanish'),
                  l1 = str_replace(l1, 'spanish, german', 'spanish'),
                  l1 = str_replace(l1, 'italian, german', 'italian'),
                  l1 = str_replace(l1, 'italian, german', 'italian'),
                  l1 = str_replace(l1, 'italia', 'italian'),                  
                  l1 = str_replace(l1, 'italiann', 'italian'),
                  l1 = str_replace(l1, 'catalan', 'spanish'), # spanish experiment
                  l1 = str_replace(l1, 'esperanto', 'spanish'), # spanish experiment
                  l1 = str_replace(l1, 'polish, silesian', 'polish'),
                  l1 = str_replace(l1, 'swedish, finnish', 'swedish'),
                  l1 = str_replace(l1, 'greek, german', 'greek'),
                  l1 = str_replace(l1, 'russian, belarusian', 'russian'),
                  l1 = str_replace(l1, 'georgian, mingrelian', 'georgian'),
                  l1 = str_replace(l1, 'hi', 'chinese'),
                  l1 = str_replace(l1, 'cchinesenese', 'chinese'),
                  l1 = str_replace(l1, 'zhuang', 'chinese')) # chinese experiment
```
-->

```{r replacements l2 info, message=FALSE, warning=FALSE}
L2_info <- mutate(L2_info,
                  l1 = str_replace(l1, 'isizulu', 'zulu'),
                  l1 = str_replace(l1, 'turkce', 'turkish'),
                  l1 = str_replace(l1, 'albanianja', 'albanian'),
                  l1 = str_replace(l1, 'swissgerman', 'german'),
                  l1 = str_replace(l1, 'austriangerman', 'german'),
                  l1 = str_replace(l1, 'español', 'spanish'),
                  l1 = str_replace(l1, 'italia', 'italian'),                  
                  l1 = str_replace(l1, 'italiann', 'italian'),
                  l1 = str_replace(l1, 'catalan', 'spanish'), # spanish experiment
                  l1 = str_replace(l1, 'esperanto', 'spanish'), # spanish experiment
                  l1 = str_replace(l1, 'hi', 'chinese'),
                  l1 = str_replace(l1, 'cchinesenese', 'chinese'),
                  l1 = str_replace(l1, 'zhuang', 'chinese')) # chinese experiment
```

## Process L2 information:

Transform NAs into "this person knows no L2":

```{r no l2 info, message=FALSE, warning=FALSE}
L2_info <- L2_info %>%
  mutate(l2 = ifelse(l2 == " " | is.na(l2), "no_L2", l2),)
```

Extract whether the person reports any English:

```{r english as l2, message=FALSE, warning=FALSE}
L2_info <- mutate(L2_info,
                  EnglishL2YesNo = str_detect(l2, 'english'))
```

Conversions:

```{r ls info cleaning, message=FALSE, warning=FALSE}
L2_info <- L2_info %>%
  mutate(l2 = str_replace_all(l2, c('potuguese' = 'portuguese',
                                    'spanich' = 'spanish',
                                    'calatan' = 'catalan',
                                    'ancient greek' = 'greek',
                                    'rumanian' = 'romanian',
                                    'estniska' = 'estonian',
                                    'calabrese' = 'italian',
                                    'dari' = 'farsi',
                                    '\\(german\\)' = 'german',
                                    'sami' = 'saami',
                                    'inari saami' = 'saami',
                                    "(a little bit )|(a little )" = "",
                                    "\\s+" = " ",
                                    "\\s+$" = "",
                                    ",(?![ ])" = ", ")))
```

## Add rhotics information:

Add new columns

```{r add rhotic cols, message=FALSE, warning=FALSE}
L2_info <- L2_info %>% 
           mutate(r_l_distinction_L1 = NA,
                  trill_real_L1 = NA,
                  trill_occ_L1 = NA,
                  r_l_distinction_L2 = NA,
                  trill_real_L2 = NA,
                  trill_occ_L2 = NA)
```

<!-- Join data for L1

```{r add rhotic info l1_old, message=FALSE, warning=FALSE, eval=FALSE}
L2_info <- L2_info %>% 
           left_join(languages_data, by = c("l1" = "Languages")) %>%
           mutate(r_l_distinction_L1 = coalesce(r_l_distinction, r_l_distinction_L1),
                  trill_real_L1 = coalesce(trill_real, trill_real_L1),
                  trill_occ_L1 = coalesce(trill_occ, trill_occ_L1)) %>%
           select(-c(glottocode, r_l_distinction, trill_real, trill_occ))

```
-->

Split and join data for L1

```{r add rhotic info l1, message=FALSE, warning=FALSE}
# Save the original 'l1' column
original_l1 <- L2_info$l1

# Run the main code
L2_info <- L2_info %>% 
  separate_rows(l1, sep = "[^[:alnum:].]+") %>% 
  left_join(languages_data, by = c("l1" = "Languages")) %>%
  group_by(ID) %>%
  mutate(r_l_distinction_L1 = if(all(is.na(r_l_distinction))) NA else max(coalesce(r_l_distinction, r_l_distinction_L1), na.rm = TRUE),
         trill_real_L1 = if(all(is.na(trill_real))) NA else max(coalesce(trill_real, trill_real_L1), na.rm = TRUE),
         trill_occ_L1 = if(all(is.na(trill_occ))) NA else max(coalesce(trill_occ, trill_occ_L1), na.rm = TRUE)) %>%
  select(-c(glottocode, r_l_distinction, trill_real, trill_occ)) %>%
  distinct(ID, .keep_all = TRUE) %>%
  ungroup()

# Replace the original 'l1' column
L2_info$l1 <- original_l1

rm(original_l1)
```

Split and join data for L2

```{r add rhotic info l2, message=FALSE, warning=FALSE}
# Save the original 'l2' column
original_l2 <- L2_info$l2

# Some replacements:
L2_info$l2 <- gsub(".", ",", L2_info$l2, fixed=TRUE)
L2_info$l2[ L2_info$l2 == "No_L2" ] <- NA

# Run the main code
L2_info <- L2_info %>% 
  separate_rows(l2, sep = "[^[:alnum:].]+") %>% 
  left_join(languages_data, by = c("l2" = "Languages")) %>%
  group_by(ID) %>%
  mutate(r_l_distinction_L2 = if(all(is.na(r_l_distinction))) NA else max(coalesce(r_l_distinction, r_l_distinction_L2), na.rm = TRUE),
         trill_real_L2 = if(all(is.na(trill_real))) NA else max(coalesce(trill_real, trill_real_L2), na.rm = TRUE),
         trill_occ_L2 = if(all(is.na(trill_occ))) NA else max(coalesce(trill_occ, trill_occ_L2), na.rm = TRUE)) %>%
  select(-c(glottocode, r_l_distinction, trill_real, trill_occ)) %>%
  distinct(ID, .keep_all = TRUE) %>%
  ungroup()

# Replace the original 'l2' column
L2_info$l2 <- original_l2

rm(original_l2)
```

Match this with the "web" data frame:

```{r l2 and web, message=FALSE, warning=FALSE}
web <- left_join(web, select(L2_info, -Language), by = c('ID' = 'ID'))
```

Rename:

```{r rename vars, message=FALSE, warning=FALSE}
web <- rename(web,
              L1_raw = l1.x,
              L2_raw = l2.x,
              L1_cleaned = l1.y,
              L2_cleaned = l2.y)
```


## Exclusions based on insufficient data:

How many data per participant?

```{r how many per participant, message=FALSE, warning=FALSE}
ppt_N <- web %>% count(ID)
```


Check whether there's anybody who doesn't have 2 responses?

```{r check response N, message=FALSE, warning=FALSE}
all(ppt_N$n == 2)
```

Which one?

```{r check who, message=FALSE, warning=FALSE}
filter(ppt_N, n != 2)
```

Get rid of these participants:

```{r remove less than 2, message=FALSE, warning=FALSE}
# Vector of participants to exclude:

excludes <- filter(ppt_N, n != 2) %>% pull(ID)

# Exclude:

web <- filter(web, !(ID %in% excludes))

rm(excludes)
```

Count data per language:

```{r counts per lang, message=FALSE, warning=FALSE}
web %>% count(Language) %>% 
  mutate(n = n / 2) %>% 
  print()
```

Not enough data for MS (only data from one speaker), as well as Tamil (only two speakers).

```{r remove too few, message=FALSE, warning=FALSE}
web <- filter(web, Language != "MS")
web <- filter(web, Language != "TA")
```

Get rid of L1 speakers that differ from the sample they are supposed to belong to:

```{r sort l1, message=FALSE, warning=FALSE}
web <- filter(web,
              !L1_cleaned %in% c('arabic',
                                 'kurdish', 'pashto',
                                 'wolof'))

# Close enough for our purposes from the perspective of kiki/bouba and the fact that language families don't cross — sorry if you are a speaker of one of these languages as we recognize the difference, but the most important thing is that these do not exert a bias from a "macro perspective" of language families:

web[web$L1_cleaned == 'dutch', ]$L1_cleaned <- 'german'
web[web$L1_cleaned == 'czech', ]$L1_cleaned <- 'polish'
web[web$L1_cleaned == 'latvian', ]$L1_cleaned <- 'polish'
```

Check which ones are misaligned:

```{r clean l1, message=FALSE, warning=FALSE}
table(web$L1_cleaned, web$Language)
web[web$L1_cleaned == 'russian', ]$Language <- 'RU'
web[web$L1_cleaned == 'armenian', ]$Language <- 'AM'
web[web$L1_cleaned == 'chinese', ]$Language <- 'CN'
web[web$L1_cleaned == 'english', ]$Language <- 'EN'
web[web$L1_cleaned == 'estonian', ]$Language <- 'EE'
web[web$L1_cleaned == 'german', ]$Language <- 'DE'
web[web$L1_cleaned == 'finnish', ]$Language <- 'FI'
web[web$L1_cleaned == 'french', ]$Language <- 'FR'
web[web$L1_cleaned == 'italian', ]$Language <- 'IT'
web[web$L1_cleaned == 'korean', ]$Language <- 'KR'
web[web$L1_cleaned == 'polish', ]$Language <- 'PL'
web[web$L1_cleaned == 'zulu', ]$Language <- 'ZU'
web[web$L1_cleaned == 'spanish', ]$Language <- 'ES'
web[web$L1_cleaned == 'swedish', ]$Language <- 'SE'
web[web$L1_cleaned == 'portuguese', ]$Language <- 'PT'
web[web$L1_cleaned == 'thai', ]$Language <- 'TH'
```

## Add language info:

Load language file:

```{r load lang file, message = FALSE}
langs <- read_csv(paste0(data, 'language_info.csv'))
```

Merge:

```{r merge lang, message=FALSE, warning=FALSE}
web <- left_join(web, langs)
```

## Create response variable

Change main response to "Resp":

```{r add resp, message=FALSE, warning=FALSE}
web <- rename(web,
               Resp = inputvalue)
```

Let's process the main predictor (the audio file shown):

```{r add condition, message=FALSE, warning=FALSE}
web <- mutate(web,
              Condition = str_replace(audio, '\\.wav', ''))
```

Create congruent/incongruent responses:

```{r congruency, message=FALSE, warning=FALSE}
web$match = web$Condition == web$Resp
```

Add trial order info (more efficient code by reviewer 1 sister study):

```{r trial, message=FALSE, warning=FALSE}
web$sound <- xfun::sans_ext(web$audio)
web$match <- web$sound == web$Resp
table(web$match)
# add trial
web$trial <- NA
for (subject in unique(web$ID)) {
  idx <- which(web$ID == subject)
  web$trial[idx] <- 1:length(idx)
}
table(web$trial)

out <- web[, c('ID', 'sex', 'Language', 'Name', 'Script', 'Family', 'Autotyp_Area', 'trial', 'sound', 'match')]

rm(idx, subject)

# Rename column names for consistency with the main analysis:

out <- rename(out, Condition = sound)

write_csv(out, paste0(data,'web_raw_trials.csv'))
```


## By-participant preprocessing

Arguably, the responses to "r" and "l", since they immediately followed each other, are not independent. 

Let's create a file of matching versus non-matching participants. First, create a vector of Condition/Resp contingency tables broken up by participant:

```{r ID tabs, message=FALSE, warning=FALSE}
ID_tabs <- with(web, table(Condition, Resp, ID))
```

Loop through this and save whether they were 100% matching:

```{r matches, message=FALSE, warning=FALSE}
matches <- numeric(dim(ID_tabs)[3])

for (i in seq_along(matches)) {
  matches[i] <- as.integer(sum(diag(ID_tabs[, , i])) == 2)
}
```

Put this together with ID info into a table:

```{r ids, message=FALSE, warning=FALSE}
ids <- unique(web$ID)
r_ppt <- tibble(ID = ids, Match = matches)

rm(ids, i, ID_tabs, matches)
```


```{r get order for r_ppt, message=FALSE, warning=FALSE}
web <- web %>%
  group_by(ID) %>%
  mutate(Order = ifelse(trial == 1 & n() == 2, 
                        ifelse(sound == "l", "l_first", "r_first"),
                        NA)) %>%
  fill(Order, .direction = "updown") %>%
  ungroup()
```

Merge this with the relevant info:

```{r merge ppt, message=FALSE, warning=FALSE}
r_ppt$Language <- web[match(r_ppt$ID, web$ID), ]$Language
r_ppt$Sex <- web[match(r_ppt$ID, web$ID), ]$sex
r_ppt$Age <- web[match(r_ppt$ID, web$ID), ]$participantage
r_ppt$Name <- web[match(r_ppt$ID, web$ID), ]$Name
r_ppt$Script <- web[match(r_ppt$ID, web$ID), ]$Script
r_ppt$Family <- web[match(r_ppt$ID, web$ID), ]$Family
r_ppt$Autotyp_Area <- web[match(r_ppt$ID, web$ID), ]$Autotyp_Area
r_ppt$L2 <- web[match(r_ppt$ID, web$ID), ]$L2_raw
r_ppt$EnglishL2YesNo <- web[match(r_ppt$ID, web$ID), ]$EnglishL2YesNo
r_ppt$Order <- web[match(r_ppt$ID, web$ID), ]$Order
r_ppt$r_l_distinction_L1 <- web[match(r_ppt$ID, web$ID), ]$r_l_distinction_L1
r_ppt$trill_real_L1 <- web[match(r_ppt$ID, web$ID), ]$trill_real_L1
r_ppt$trill_occ_L1 <- web[match(r_ppt$ID, web$ID), ]$trill_occ_L1
r_ppt$r_l_distinction_L2 <- web[match(r_ppt$ID, web$ID), ]$r_l_distinction_L2
r_ppt$trill_real_L2 <- web[match(r_ppt$ID, web$ID), ]$trill_real_L2
r_ppt$trill_occ_L2 <- web[match(r_ppt$ID, web$ID), ]$trill_occ_L2
```

Write to file:

```{r save cleaned data, message=FALSE, warning=FALSE}
write_csv(r_ppt, paste0(data, 'web_experiment_cleaned.csv'))

web <- read_csv(paste0(data, 'web_experiment_cleaned.csv'))
```

Remove unnecessary objects

```{r remove objects web, message=FALSE, warning=FALSE}
rm(L2_info, langs, out, ppt_N, r_ppt)
```

# Field experiment

```{r load demographics field, message = FALSE}
# Load:
demographics <- read_delim(paste0(data, 'fieldwork-personal-data.csv'), delim = ';')
```

## Data cleaning & exclusions

Fix last column name:

```{r fix colname, message=FALSE, warning=FALSE}
colnames(field)[ncol(field)] <- 'PA08'
```

The field response file is wide format. Let's transform this to long format:

```{r reshape table, message=FALSE, warning=FALSE}
# Remove rows with names "bouba-ort", "bouba-aud", "kiki-ort", and "kiki-aud"
field <- field %>%
  filter(!file %in% c("bouba-ort", "bouba-aud", "kiki-ort", "kiki-aud"))

#field[1, -1] <- gsub("rock", "r", field[1, -1])

field <- field %>%
  select(-file) %>%
  pivot_longer(everything(), names_to = "ID", values_to = "value") %>%
  arrange(ID) %>%  # Arrange by ID
  group_by(ID) %>%
  mutate(audio = rep(c("r", "l"), each = n() / 2),  # Repeat "r" and "l" for each ID
         response = ifelse(value %in% c("0", "NA", " "), NA, value)) %>%
  ungroup() %>%
  select(-value)  # Remove the original "value" column
```

Get the language info out of the ID column:

```{r get lang info, message=FALSE, warning=FALSE}
field <- mutate(field,
                Language = str_extract(ID, '[A-Z]+'),
                Participant = str_extract(ID, '\\d+'))
```

Add age and gender info:

```{r add age gender, message=FALSE, warning=FALSE}
# Rename so that key columns match names:

demographics <- rename(demographics,
                       ID = subject,
                       Age = age,
                       Sex = gender,
                       L2 = "other-lang")

# Get only that what is needed:

demographics <- select(demographics,
                       ID, Age, Sex, L2)

# Join:

field <- left_join(field, demographics)

```

Discard Polish (only one data point) and US (no data collected for this part of the experiment):

```{r exclude polish, message=FALSE, warning=FALSE}
field <- filter(field, Language != 'US')
field <- filter(field, Language != 'PL')
```

How many before exclusions:

```{r how many field, message=FALSE, warning=FALSE}
length(unique(field$ID))
```

Exclude participants under 18:

```{r exclude field, message=FALSE, warning=FALSE}
field <- filter(field, Age >= 18)
```

Check if there is anyone who doesn't have two any responses.

```{r, message=FALSE, warning=FALSE}
field %>%
  group_by(ID) %>%
  summarise(AllResponsesEmptyOrNA = all(is.na(response) | response == "")) %>%
  filter(AllResponsesEmptyOrNA)
```

Remove this participant.

```{r, message=FALSE, warning=FALSE}
field <- field %>%
  filter(ID != "BE40")
```


## Process L1 info:

Add language name information to the data frame:

```{r add info on contrast, message=FALSE, warning=FALSE}
field$Name <- NA
field[field$Language == 'BE', ]$Name <- 'english'
field[field$Language == 'DE', ]$Name <- 'german'
field[field$Language == 'SR', ]$Name <- 'portuguese'
field[field$Language == 'VA', ]$Name <- 'daakie'
field[field$Language == 'BR', ]$Name <- 'berber'
field[field$Language == 'PA', ]$Name <- 'palikur'
```

## Process L2 info:

Create a table with one data point per participant with L2 information:

```{r process l2, message=FALSE, warning=FALSE}
L2_info_field <- filter(field, !duplicated(ID)) %>% select(ID, Language, Sex, Name, L2)
```

Transform NAs into "this person knows no L2":

```{r add no l2, message=FALSE, warning=FALSE}
L2_info_field <- L2_info_field %>% 
  mutate(L2 = ifelse(L2 %in% c('0', 'NA', ' '), NA, L2)) %>% 
  mutate(L2 = ifelse(is.na(L2), 'no_L2', L2))

# also do it in the field df for the future
field <- field %>% 
  mutate(L2 = ifelse(L2 %in% c('0', 'NA', ' '), NA, L2))
```

Extract whether the person reports any English:

```{r add english, message=FALSE, warning=FALSE}
L2_info_field <- mutate(L2_info_field,
                  EnglishL2YesNo = ifelse(grepl('english', L2, ignore.case = TRUE), TRUE, FALSE))
```

Conversions:

```{r l2 conversions, message=FALSE, warning=FALSE}
L2_info_field <- L2_info_field %>%
  mutate(L2 = str_replace_all(L2, regex('British Sign Language', ignore_case = TRUE), 'BSL'),
         L2 = str_replace_all(L2, regex('Mandarin', ignore_case = TRUE), 'chinese'),
         L2 = str_replace_all(L2, regex('Moroccan Arabic', ignore_case = TRUE), 'arabic'),
         L2 = str_replace_all(L2, regex('Standard Arabic', ignore_case = TRUE), 'arabic'))


L2_info_field <- L2_info_field %>%
  mutate(L2 = ifelse(L2 %in% c("no_L2", "BSL"), L2, str_to_lower(L2))) %>%
  rowwise() %>%
  mutate(L2 = paste(unique(unlist(str_split(L2, ",\\s*"))), collapse = ", "))

L2_info_field$L2 <- sub("\\s*\\(restricted reading skills\\)", "", L2_info_field$L2)
```

## Add rhotics information:

```{r add rhotic cols field, message=FALSE, warning=FALSE}
L2_info_field <- L2_info_field %>% 
           mutate(r_l_distinction_L1 = NA,
                  trill_real_L1 = NA,
                  trill_occ_L1 = NA,
                  r_l_distinction_L2 = NA,
                  trill_real_L2 = NA,
                  trill_occ_L2 = NA)
```

Join data for L1

```{r add rhotic info l1 field, message=FALSE, warning=FALSE}
L2_info_field <- L2_info_field %>% 
           left_join(languages_data, by = c("Name" = "Languages")) %>%
           mutate(r_l_distinction_L1 = coalesce(r_l_distinction, r_l_distinction_L1),
                  trill_real_L1 = coalesce(trill_real, trill_real_L1),
                  trill_occ_L1 = coalesce(trill_occ, trill_occ_L1)) %>%
           select(-c(glottocode, r_l_distinction, trill_real, trill_occ))

```

Split and join data for L2

```{r add rhotic info l2 field, message=FALSE, warning=FALSE}
# Save the original 'l2' column
original_L2 <- L2_info_field$L2

# Run the main code
L2_info_field <- L2_info_field %>% 
  separate_rows(L2, sep = ",\\s*") %>% 
  left_join(languages_data, by = c("L2" = "Languages")) %>%
  group_by(ID) %>%
  mutate(r_l_distinction_L2 = if(all(is.na(r_l_distinction))) NA else max(coalesce(r_l_distinction, r_l_distinction_L2), na.rm = TRUE),
         trill_real_L2 = if(all(is.na(trill_real))) NA else max(coalesce(trill_real, trill_real_L2), na.rm = TRUE),
         trill_occ_L2 = if(all(is.na(trill_occ))) NA else max(coalesce(trill_occ, trill_occ_L2), na.rm = TRUE)) %>%
  select(-c(glottocode, r_l_distinction, trill_real, trill_occ)) %>%
  distinct(ID, .keep_all = TRUE) %>%
  ungroup()

# Replace the original 'L2' column
L2_info_field$L2 <- original_L2

rm(original_L2)
```

Match this with the "field" data frame:

```{r l2 and field, message=FALSE, warning=FALSE}
field <- left_join(field, select(L2_info_field, -Language), by = c('ID' = 'ID'))
```

Rename:

```{r rename vars field, message=FALSE, warning=FALSE}
field <- field %>% 
  rename(Name = Name.x,
         L2_raw = L2.x,
         L2_cleaned = L2.y) %>% 
  select(-Name.y)
```

## Exclusions and descriptive info

How many data per participant?

```{r how many per participant field, message=FALSE, warning=FALSE}
ppt_N <- field %>% count(ID)
```

Check whether there's anybody who doesn't have 2 responses?

```{r check response N field, message=FALSE, warning=FALSE}
all(ppt_N$n == 2)
```

All have two responses.

```{r remove less than 2 field, message=FALSE, warning=FALSE}
rm(ppt_N)
```

Count data per language:

```{r counts per lang field, message=FALSE, warning=FALSE}
field %>% count(Name) %>% 
  mutate(n = n / 2) %>% 
  print()
```

Palikur has the fewest, but still ok.

## Add language info:

Load language file:

```{r load lang file field, message = FALSE, message=FALSE, warning=FALSE}
langs <- read_csv(paste0(data, 'language_info_field.csv'))
```

Merge:

```{r merge lang field, message=FALSE, warning=FALSE}
field <- left_join(field, langs, by = "Language")

field <- field %>% 
  rename(Name = Name.y) %>% 
  select(-Name.x) %>% 
  rename(Sex = Sex.y) %>% 
  select(-Sex.x)
```

## Create response variable

Create congruent/incongruent responses:

```{r congruency field, message=FALSE, warning=FALSE}
field$match = field$audio == field$response
```

<!--If this results in NA, put FALSE there (working against our hypothesis).

```{r matches field_old, message=FALSE, warning=FALSE, eval=FALSE}
field <- field %>%
  mutate(match = ifelse(is.na(match), FALSE, match))
```
-->

There are `r sum(is.na(field$match))` with `NA` for the match: participant DE08 matched both lines with the sound /l/ and participant DE09 did not provide an answer for the sound /r/ -- we chose to remove these participants from the data because arguably they failed to follow the task instructions correctly.

```{r matches field, message=FALSE, warning=FALSE}
field <- field[ !is.na(field$match), ]
```

The task here was a little different, the matching was done simultaneously, therefore, there is no information on the order.

```{r create output field, message=FALSE, warning=FALSE}
out <- field[, c('ID', 'Sex', 'Language', 'Name', 'Script', 'Family', 'Autotyp_Area', 'audio', 'match')]

# Rename column names for consistency with the main analysis:

out <- rename(out, Condition = audio)

write_csv(out, paste0(data,'field_raw_trials.csv'))
```

## By-participant preprocessing

Arguably, the responses to "r" and "l", since they were presented together, are not independent of each other.

Let's create a data.frame for each participant whether they were 100% matching.

```{r ids field, message=FALSE, warning=FALSE}
r_ppt <- field %>%
  group_by(ID) %>%
  summarise(Match = as.integer(all(match == TRUE)))
```

Merge this with the relevant info:

```{r merge ppt field, message=FALSE, warning=FALSE}
r_ppt$Language <- field[match(r_ppt$ID, field$ID), ]$Language
r_ppt$Sex <- field[match(r_ppt$ID, field$ID), ]$Sex
r_ppt$Age <- field[match(r_ppt$ID, field$ID), ]$Age
r_ppt$Name <- field[match(r_ppt$ID, field$ID), ]$Name
r_ppt$Script <- field[match(r_ppt$ID, field$ID), ]$Script
r_ppt$Family <- field[match(r_ppt$ID, field$ID), ]$Family
r_ppt$Autotyp_Area <- field[match(r_ppt$ID, field$ID), ]$Autotyp_Area
r_ppt$L2 <- field[match(r_ppt$ID, field$ID), ]$L2_raw
r_ppt$EnglishL2YesNo <- field[match(r_ppt$ID, field$ID), ]$EnglishL2YesNo
r_ppt$r_l_distinction_L1 <- field[match(r_ppt$ID, field$ID), ]$r_l_distinction_L1
r_ppt$trill_real_L1 <- field[match(r_ppt$ID, field$ID), ]$trill_real_L1
r_ppt$trill_occ_L1 <- field[match(r_ppt$ID, field$ID), ]$trill_occ_L1
r_ppt$r_l_distinction_L2 <- field[match(r_ppt$ID, field$ID), ]$r_l_distinction_L2
r_ppt$trill_real_L2 <- field[match(r_ppt$ID, field$ID), ]$trill_real_L2
r_ppt$trill_occ_L2 <- field[match(r_ppt$ID, field$ID), ]$trill_occ_L2
```

Write to file:

```{r save cleaned data field, message=FALSE, warning=FALSE}
write_csv(r_ppt, paste0(data, 'field_experiment_cleaned.csv'))

field <- read_csv(paste0(data, 'field_experiment_cleaned.csv'))
```

Remove unnecessary objects

```{r remove objects field, message=FALSE, warning=FALSE}
rm(L2_info_field, langs, out, r_ppt, demographics)
```

