---
title: "R/L across languages: Modeling"
author: "Aleksandra Ćwiek"
date: "2023-10-26"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 4
    toc_float: yes
    df_print: paged
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: '3'
  html_notebook:
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Introduction:

This script uses the output of "r_l_preparation.Rmd" and returns all values reported in the paper.

# Setup:

Load packages:

```{r, message = FALSE, warning = FALSE}
library(tidyverse) # data processing
library(brms) # bayesian models
library(cmdstanr)
library(ggdist) # for plotting
# option for Bayesian regression models:
# use all available cores for parallel computing
options(mc.cores = parallel::detectCores())

# Set the script's path as working directory
parentfolder = rstudioapi::getActiveDocumentContext()$path 
setwd(dirname(parentfolder))
parentfolder <- getwd()

models        <- paste0(parentfolder, '/models/')
plots         <- paste0(parentfolder, '/plots/')
data          <- paste0(parentfolder, '/data/')
```

For reproducible reporting:

```{r, message = FALSE, warning = FALSE}
packageVersion('tidyverse')
packageVersion('brms')
packageVersion('cmdstanr')
packageVersion('ggdist')
```

Load ggplot2 theme:

```{r}
source('theme_timo.R')
```

Load data:

```{r, message = FALSE, warning = FALSE}
web <- read_csv(paste0(data, 'web_experiment_cleaned.csv'))
web_raw <- read_csv(paste0(data, '/web_raw_trials.csv'))
```

# Online experiment

## Descriptive statistics

First, how many participants?

```{r}
nrow(web)
```

First, how many languages?

```{r}
web %>% count(Language) %>% nrow()
```

How many families?

```{r}
web %>% count(Family) %>% nrow()
```


How many non-R/L-distinct L1 among the languages?

```{r}
web %>% count(Language, L1_distinction) %>% count(L1_distinction)
```


What is the grand average congruent behavior?

```{r}
mean(web$Match)
```

87%

What about only among those who have L1 without the distinction?

```{r}
web %>%
  filter(L1_distinction == "no") %>%
  summarize(mean_match = mean(Match, na.rm = TRUE))
```

84%

What about only among those who have L1 without the distinction and no L2 that distinguishes?

```{r}
web %>%
  filter(L1_distinction == 'no') %>%
  filter(!EnglishL2YesNo) %>% 
  filter(no_dist_L2 == 'FALSE') %>% 
  summarize(mean_match = mean(Match, na.rm = TRUE))
```

Also 84%

Compute average matching behavior per language and sort:

```{r}
web_avg <- web %>%
  group_by(Language) %>% 
  summarize(M = mean(Match)) %>% 
  arrange(desc(M)) %>% 
  mutate(percent = round(M, 2) * 100,
         percent = str_c(percent, '%'))

# Show:

web_avg %>% print(n = Inf)
```

Check some demographics, also to report in the paper. First, the number of participants per language:

```{r}
web %>% 
  count(Name, sort = TRUE) %>% print(n = Inf)
```


Then, the number of L1 speakers who have R/L distinction vs. who don't (we can think if we put the approximant and trill together):

```{r}
web %>% count(L1_distinction) %>%
  mutate(prop = n / sum(n),
         percent = round(prop, 2) * 100,
         percent = str_c(percent, '%'))
```

How many people are not bilinguals?

```{r}
sum(is.na(web$L2)) / nrow(web)

# bilinguals:

1 - sum(is.na(web$L2)) / nrow(web)
```

Check how many people knew English as their L2:

```{r}
web %>% count(EnglishL2YesNo) %>% 
  mutate(prop = n / sum(n),
         percent = round(prop, 2) * 100,
         percent = str_c(percent, '%'))
```

Of those that don't use a R/L distinction in their L1, how many use R/L distinction in their L2? (double-check if logic alright!)

```{r}
web %>%
  filter(L1_distinction == 'no') %>% 
  count(no_dist_L2 == FALSE) %>% 
  mutate(prop = n / sum(n),
         percent = round(prop, 2) * 100,
         percent = str_c(percent, '%'))
```

How many "pure" speakers were there?, i.e., those people that 1) don't know English, 2) don't use an L1 with a R/L distinction, and 3) don't know an L2 that distinguishes R/L.

```{r}
web %>% 
  filter(L1_distinction == 'no') %>%  # excludes English as well
  filter(!EnglishL2YesNo) %>% 
  filter(no_dist_L2 == 'FALSE') %>% 
  nrow()
```

43 people.

Let's check if this is correct. This gives the list of all participants for whom this applies.

```{r}
web %>% 
    filter(L1_distinction == 'no' & !EnglishL2YesNo & no_dist_L2 == FALSE) %>% 
    print(n = 50)
```

Are these really "pure"? What languages do they speak?

```{r}
web %>% 
  filter(L1_distinction == 'no') %>%  # excludes English as well
  filter(!EnglishL2YesNo) %>% 
  filter(no_dist_L2 == 'FALSE') %>% 
  count(Language)
```

Mostly Chinese and Japanese.


Nevertheless, let's explore whether these also show kiki/bouba matches?

```{r}
web %>% 
  filter(L1_distinction == 'no') %>%  # excludes English as well
  filter(!EnglishL2YesNo) %>% 
  filter(no_dist_L2 == 'FALSE') %>% 
  count(Match) %>% 
  mutate(prop = n / sum(n),
         percent = round(prop, 2) * 100,
         percent = str_c(percent, '%'))
```

Yes, as shown above 84%.

## Main model:


Set parameters that will be used for all MCMC sampling:

```{r}
mywarmup <- 4000
myiter <- 6000
```

Check the distribution of scripts across families to make decisions about random effects structure:

```{r}
table(web$Family, web$L1_distinction)
```


Let's factor code the Order and L1_distinction effects:

```{r}
web <- mutate(web,
               Order = factor(Order, levels = c('r_first', 'l_first')),
               L1_distinction = factor(L1_distinction, levels = c('no', 'approximant', 'trill')))
```

Check that the order effect is approximately balanced:

```{r}
web %>% count(Order) %>%
  mutate(prop = n / sum(n))
```

I omit some things for now... Because for now, L1_distinction is three-way.

```{r}
web <- mutate(web,
               order_num = ifelse(Order == 'r_first', -0.5, +0.5))
```

The main model with Order and L1_distinction fixed effects, as well as Language and Family random effects.

```{r, message = FALSE, warning = FALSE}
web_mdl <- brm(Match ~ 1 + order_num + L1_distinction +
                 (1 + order_num|Language) + 
                 (1 + order_num|Family),
                data = web,
                family = bernoulli(link = 'logit'),
                backend = "cmdstanr",
                seed = 42,
                cores = 4,
                iter = 2000,
                warmup = 1000,
                control = list(adapt_delta = 0.995,
                               max_treedepth = 13),
                file = paste0(models, "web_mdl.rds"))
```

Show the model:

```{r}
web_mdl
```



